---
output: pdf_document
---

\chapter{Results 1 - Descriptive statistics}\label{ch:results-1}
\thispagestyle{empty}

This chapter outlines the characteristics of the cohort and distributions of pertinent variables in three sections. The first section contains tabulation and visualisation of counts and percentages of individuals in the cohort overall, by age & sex, by mortality, and by geography.  The next section focusses on the proxy measurement of multimorbidity, repeat medicine counts, and its distribution by sociodemographic variables. The final section focusses on deprivation. A table with counts and percentages split by SIMD deprivation quintile is provided plus visualisations of the distribution of SIMD deciles by age & sex and local authority. Where distributions were found to be similar across all years of data a reference year, financial year 2013/14, is used for visualisation purposes. This is the middle year of the study period. 

Descriptive statistics of social care and unscheduled care measures are not presented in this chapter but in chapters \ref{ch:res2} and \ref{ch:res-3} respectively where the focus is on these variables specifically. 

\section{Age, sex, mortality, and geography}\label{sec:results1-descr}

\subsection{Participants}\label{subsec:results1-particp}

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{threeparttable}
\begin{tabular}{@{}llrrrrr@{}}
\toprule
\textbf{Variable} & \textbf{Value} & \textbf{2011/12} & \textbf{2012/13} & \textbf{2013/14} & \textbf{2014/15} & \textbf{2015/16} \\ \midrule
\textbf{N in cohort} &  & 870923 & 895113 & 908300 & 920647 & 924323 \\
\textbf{N died} &  & 35874 (4.1) & 41702 (4.7) & 39346 (4.3) & 42365 (4.6) & 37992 (4.1) \\
\textbf{Sex} & Male & 382598 (43.9) & 395927 (44.2) & 404478 (44.5) & 412128 (44.8) & 415968 (45.0) \\
\textbf{} & Female & 488325 (56.1) & 499186 (55.8) & 503822 (55.5) & 508519 (55.2) & 508355 (55.0) \\
\textbf{Age (years)} & Mean (SD) & 74.8 (7.4) & 74.7 (7.4) & 74.7 (7.4) & 74.7 (7.4) & 74.7 (7.4) \\
\textbf{Age group (years)} & 65-69 & 256764 (29.5) & 269044 (30.1) & 274946 (30.3) & 278708 (30.3) & 282563 (30.6) \\
\textbf{} & 70-74 & 206133 (23.7) & 208831 (23.3) & 211466 (23.3) & 214391 (23.3) & 215895 (23.4) \\
\textbf{} & 75-79 & 170455 (19.6) & 173475 (19.4) & 175402 (19.3) & 176749 (19.2) & 175279 (19.0) \\
\textbf{} & 80-84 & 124680 (14.3) & 127608 (14.3) & 128500 (14.1) & 130149 (14.1) & 130135 (14.1) \\
\textbf{} & 85-89 & 73705 (8.5) & 74938 (8.4) & 75992 (8.4) & 77620 (8.4) & 77694 (8.4) \\
\textbf{} & 90-94 & 30737 (3.5) & 32712 (3.7) & 33711 (3.7) & 34149 (3.7) & 33428 (3.6) \\
\textbf{} & 95 plus & 8449 (1.0) & 8505 (1.0) & 8283 (0.9) & 8881 (1.0) & 9329 (1.0) \\
\textbf{N with  \textgreater{}=1 repeat medicine} &  & 775647 (89.1) & 795502 (88.9) & 807927 (88.9) & 818717 (88.9) & 821862 (88.9) \\
\textbf{Receiving social care} &  & 88036 (10.1) & 95827 (10.7) & 95128 (10.1) & 106242 (11.1) & 107441 (11.6) \\
\textbf{Receiving home care} &  & 41642 (4.8) & 42149 (4.7) & 42746 (4.7) & 42232 (4.6) & 41172 (4.5) \\
\textbf{Resident in care home} &  & 36514 (4.4) & 33293 (4.0) & 31397 (3.7) & 29107 (3.3) & 28156 (3.2) \\
\textbf{\textgreater{}= 1 USC episode} &  & 259147 (29.8) & 272580 (30.5) & 270874 (29.8) & 289539 (31.4) & 285059 (30.8) \\
\textbf{\textgreater{}=1 hospital admission} &  & 133158 (15.3) & 141873 (15.8) & 141746 (15.6) & 147779 (16.1) & 145770 (15.8) \\
\textbf{\textgreater{}=1 A \& E attendance} &  & 172754 (19.8) & 181732 (20.3) & 184101 (20.3) & 190588 (20.7) & 185767 (20.1) \\ \bottomrule
\end{tabular}%
\begin{tablenotes}
\item All values n(\%) unless otherwise stated 
\item \emph{Column}-wise percentages within grouped variables
\end{tablenotes}
\end{threeparttable}%
}
\caption{Characteristics of study population}
\label{tab:results1-cohorttab}
\end{table}

There were 1,100,675 individuals included within the study across all years with increasing numbers in each financial year as shown in Table \ref{tab:results1-cohorttab}.  Percentages in each gender group and of those that died were relatively stable across the study period. Age was also relatively stable across all years with almost one-third of the cohort in each year being aged between 65 and 69 years old whilst over half were under the age of 75. 

In terms of healthcare, the vast majority of individuals had a least one repeat prescription in any given year, whilst the percentage of individuals using unscheduled care services (including unplanned admissions and A & E attendances) were relatively stable over time. In contrast, the pattern in \emph{absoloute} numbers with a prescription or using health services shows a gradual increase over time shadowing the increase in overall numbers in the cohort in each year.  

Social care services show slightly less stability in terms of numbers and percentages of individuals receiving care. A jump of approximately 10,000 people receiving any form of social care is found between financial years 2013/14 and 2014/15. This pattern is not observed for those receiving home care services where absolute numbers are relatively stable across the study period. Conversely, the number flagged as living in a care home shows a gradual decrease over time.   

\FloatBarrier
\subsection{Age and sex}\label{subsec:results1-age-sex}

\begin{figure}[]
  \centering
    \includegraphics{figures/chapter-results-1/02_age_sex_plot.png}
    \caption{Age and sex distribution. N = 924,323}
    \label{fig:results1-age-sex}
\end{figure}

Using financial year 2015/16 as a reference, figure \ref{fig:results1-age-sex} shows the breakdown by age and sex (similar patterns are seen across all years of data). There are more females than males in every age group with the difference between sexes widening as age increases. Over half of the cohort are under the age of 74 and over a quarter are between the ages of 75 and 84. 

\FloatBarrier
\subsection{Mortality}\label{subsec:results1-mortality}

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-1/03_deaths_plot.png}
    \caption{Mortality. N = 37,992}
    \label{fig:results1-mortality}
\end{figure}

In 2015/16, 4.1% (37,992 individuals) of the cohort died. Figure \ref{fig:results1-mortality} shows how these deaths were distributed across age and sex groups (with similar patterns seen in other years). The highest percentages of deaths (approximately 20%) occurred in the 80-84 and 85-89 age groups with females accounting for over half of these. The difference in the percentage of deaths between males and females increases from the 80-84 age groups to older age groups, whereas higher percentages of men died in age groups below 80-84.


\FloatBarrier
\subsection{Local authorities}\label{subsec:results1-la}

Figure \ref{fig:results1-la} shows the percentage of individuals in the cohort that lived in each local authority area in financial year 2015/16. The largest percentages of individuals are found in Glasgow and Edinburgh city areas, whilst the smallest number are found in the three island local authorities. This is as expected based on the general population distribution. 

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-1/04_la_count.png}
    \caption{Distribution of cohort by local authority. N = 924,323}
    \label{fig:results1-la}
\end{figure}

```{r MM section, echo=FALSE,eval=FALSE}
##################################################################################################
```
\FloatBarrier
\section{Repeat medicines}\label{sec:results1-mm-meds}
\subsection{Distribution by sex}\label{subsec:results1-meds-sex}

The distribution of number of repeat medicines prescribed in 2015/16 reveals the most common value to be zero (figure \ref{fig:results1-meds-hist}). Otherwise, a positively-skewed normal distribution is seen with higher numbers of females receiving every number of repeat medicines. This pattern is repeated across all study years.     

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-1/10_total_meds_hist_2.png}
    \caption{Number of individuals receiving specific counts of repeat medicines. N = 908,295}
    \label{fig:results1-meds-hist}
\end{figure}

\FloatBarrier
\subsection{Percentage by sex, age group, and deprivation}\label{subsec:results1-meds-age}

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-1/12_meds_age_count_plot.png}
    \caption{Repeat medicine proportion by age group}
    \label{fig:results1-meds-age-grp}
\end{figure}

Despite overall numbers for specific counts of medicines being higher for females, the \emph{percentage} of males and females receiving repeat medicines is similar as shown in the left-hand panel of figure \ref{fig:results1-meds-age-grp} **fig needs changed**. There are, however, differences in the pattern of percentages by age groups and SIMD deciles. The shape of the distribution shifts further to the right for both older age groups and more deprived SIMD deciles. Combining these factors (figure !!) reveals that the differences in deprivation deciles is more pronounced at younger ages and is much less obvious in older age groups where overall numbers are smaller.    

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-1/11_meds_simd_count_plot.png}
    \caption{Repeat medicine proportion by SIMD decile}
    \label{fig:results1-meds-simd}
\end{figure}

\FloatBarrier
\subsection{Multimorbidity (repeat medicine) groups}\label{subsec:results1-meds}

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-1/08_pis_group_plot.png}
    \caption{Distribution of repeat medicines by group (labelling needs changed)}
    \label{fig:results1-meds-grp}
\end{figure}

Figure \ref{fig:results1-meds-grp} shows the distribution of the percentage of individuals receiving numbers of repeat medicines within defined groups. Each group contains either slightly more or less than 25% of the total cohort with 0-2 and 3-5 groups being slightly larger. The same pattern is seen across all years of data. These groups are used as a proxy measurement of multimorbidity in models reported in chapters \ref{ch:res2} and \ref{ch:res-3}. 

\begin{table}[]
\centering
\resizebox{\textwidth}{!}{%
\begin{threeparttable}
\begin{tabular}{@{}llrrrrr@{}}
\toprule
\textbf{Variable} & \textbf{Value} & \textbf{\begin{tabular}[c]{@{}r@{}}Repeat Meds:\\ 0-2\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Repeat Meds:\\ 3-5\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Repeat Meds:\\ 6-8\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Repeat Meds:\\ 9+\end{tabular}} & \textbf{Total} \\ \midrule
\textbf{Sex} & Male & 117562 (28.3) & 119689 (28.8) & 94790 (22.8) & 83927 (20.2) & 415968 \\
\textbf{} & Female & 132807 (26.1) & 141500 (27.8) & 116330 (22.9) & 117718 (23.2) & 508355 \\
\textbf{Mortality} & Dead & 8326 (21.9) & 9182 (24.2) & 8657 (22.8) & 11827 (31.1) & 37992 \\
\textbf{} & Alive & 242043 (27.3) & 252007 (28.4) & 202463 (22.8) & 189818 (21.4) & 886331 \\
\textbf{Age} & Mean (SD) & 72 (6.3) & 74.4 (7.2) & 76.3 (7.6) & 77 (7.5) &  \\
\textbf{Age group} & 65-69 & 113437 (40.1) & 81178 (28.7) & 48460 (17.2) & 39488 (14.0) & 282563 \\
\textbf{} & 70-74 & 63704 (29.5) & 64409 (29.8) & 46413 (21.5) & 41369 (19.2) & 215895 \\
\textbf{} & 75-79 & 36591 (20.9) & 50161 (28.6) & 44150 (25.2) & 44377 (25.3) & 175279 \\
\textbf{} & 80-84 & 20398 (15.7) & 34397 (26.4) & 36557 (28.1) & 38783 (29.8) & 130135 \\
\textbf{} & 85-89 & 10337 (13.3) & 19841 (25.5) & 22793 (29.3) & 24723 (31.8) & 77694 \\
\textbf{} & 90-94 & 4479 (13.4) & 8612 (25.8) & 10033 (30.0) & 10304 (30.8) & 33428 \\
\textbf{} & 95 plus & 1423 (15.3) & 2591 (27.8) & 2714 (29.1) & 2601 (27.9) & 9329 \\
\textbf{SIMD Decile} & 1 & 14317 (18.6) & 18460 (24.0) & 19313 (25.1) & 24980 (32.4) & 77070 \\
\textbf{} & 2 & 17247 (20.2) & 21936 (25.7) & 21373 (25.0) & 24786 (29.0) & 85342 \\
\textbf{} & 3 & 20252 (21.3) & 25209 (26.6) & 23726 (25.0) & 25734 (27.1) & 94921 \\
\textbf{} & 4 & 23053 (23.9) & 26627 (27.7) & 23088 (24.0) & 23532 (24.4) & 96300 \\
\textbf{} & 5 & 25748 (26.1) & 27977 (28.4) & 22904 (23.3) & 21834 (22.2) & 98463 \\
\textbf{} & 6 & 28867 (28.9) & 28914 (29.0) & 22146 (22.2) & 19940 (20.0) & 99867 \\
\textbf{} & 7 & 29569 (30.9) & 27851 (29.1) & 20850 (21.8) & 17415 (18.2) & 95685 \\
\textbf{} & 8 & 28977 (31.3) & 27783 (30.0) & 19723 (21.3) & 15981 (17.3) & 92464 \\
\textbf{} & 9 & 30020 (32.3) & 28256 (30.4) & 19781 (21.3) & 15005 (16.1) & 93062 \\
\textbf{} & 10 & 32319 (35.5) & 28176 (30.9) & 18216 (20.0) & 12438 (13.6) & 91149 \\
\textbf{Any form of Social Care} & No Social Care & 241929 (29.6) & 240595 (29.5) & 181094 (22.2) & 153264 (18.8) & 816882 \\
\textbf{} & Social Care & 8440 (7.9) & 20594 (19.2) & 30026 (27.9) & 48381 (45.0) & 107441 \\
\textbf{Any USC episode} & No USC & 209153 (32.7) & 196097 (30.7) & 137476 (21.5) & 96538 (15.1) & 639264 \\
\textbf{} & USC & 41216 (14.5) & 65092 (22.8) & 73644 (25.8) & 105107 (36.9) & 285059 \\ \bottomrule
\end{tabular}%
\begin{tablenotes}
\item All values n(\%) unless otherwise stated 
\item \emph{Row}-wise percentages within grouped variables
\item USC = Unscheduled care
\end{tablenotes}
\end{threeparttable}%
}
\caption{Cohort 2013/14 by Multimorbidity (repeat medicine) groups}
\label{tab:res1-mmgroups}
\end{table}


A higher percentage of females are found in the highest multimorbidity group (9+ repeat medicines), whereas a higher percentage of men is found in the two lowest groups(0-2 & 3-5 repeat medicines, table \ref{tab:res1-mmgroups}). The average age of individuals in each multimorbidity group rises with severity from 72 years in the lowest group to 77 years in the highest. Percetntages of indiviudals in the lowest multimorbidity group run from high to low thorugh age groups whilst the opposite is found in the highest multimorbidiy group.   

A stark pattern is seen in deprivation status. Less than one-in-five individuals living in decile 1 received 0-2 repeat medicines whereas almost one-third of those in decile 10 are in this lowest multimorbidity group. The percentage of individuals in the 0-2 group increases in a linear fashion through the deprivation deciles from most deprived to most affluent. Conversely, almost a third (30.8%) of those in the most deprived SIMD decile were in the highest multimorbidity group. Only 13.6% of those in the most affluent decile were found in this group. Again, a (negative) linear trend in percentages going through SIMD deciles is observed. 

There is a noticeable relationship between service use and multimorbidity groups. Only 7.6% of those receiving social care and 14.9% of those having at least one unscheduled care episode received between 0-2 repeat medicines. This compared to 44.9% of those receiving social care and 36.9% of those using unscheduled care services who were in the 9 and over repeat medicine group. 

```{r DEPRIVATION, eval=FALSE, echo=FALSE}
##################################################################################################
```

\FloatBarrier
\section{Deprivation}\label{sec:results1-deprivation}


\subsection{Participants}\label{subsec:dep-parps}

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{threeparttable}
\begin{tabular}{@{}llrrrrrr@{}}
\toprule
\textbf{Variable} & \textbf{Value} & \textbf{\begin{tabular}[c]{@{}r@{}}Quintile 1\\ (most deprived)\end{tabular}} & \textbf{Quintile 2} & \textbf{Quintile 3} & \textbf{Quintile 4} & \textbf{\begin{tabular}[c]{@{}r@{}}Quintile 5\\ (most affluent)\end{tabular}} & \textbf{Total} \\ \midrule
\textbf{Sex} & Male & 71081 (17.1) & 83977 (20.2) & 89996 (21.6) & 86628 (20.8) & 84286 (20.3) & 415968 \\
\textbf{} & Female & 91331 (18.0) & 107244 (21.1) & 108334 (21.3) & 101521 (20.0) & 99925 (19.7) & 508355 \\
\textbf{Mortality} & Died & 8398 (22.1) & 8688 (22.9) & 8099 (21.3) & 6848 (18.0) & 5959 (15.7) & 37992 \\
\textbf{} & Alive & 154014 (17.4) & 182533 (20.6) & 190231 (21.5) & 181301 (20.5) & 178252 (20.1) & 886331 \\
\textbf{Age} & Mean (SD) & 74.7 (7.3) & 74.9 (7.4) & 74.7 (7.4) & 74.6 (7.4) & 74.7 (7.5) &  \\
\textbf{Age group} & 65-69 & 48977 (17.3) & 56371 (19.9) & 60557 (21.4) & 59224 (21.0) & 57434 (20.3) & 282563 \\
\textbf{} & 70-74 & 37526 (17.4) & 44042 (20.4) & 46823 (21.7) & 44840 (20.8) & 42664 (19.8) & 215895 \\
\textbf{} & 75-79 & 31883 (18.2) & 37192 (21.2) & 37658 (21.5) & 34598 (19.7) & 33948 (19.4) & 175279 \\
\textbf{} & 80-84 & 23666 (18.2) & 28274 (21.7) & 27279 (21.0) & 25350 (19.5) & 25566 (19.6) & 130135 \\
\textbf{} & 85-89 & 13338 (17.2) & 16545 (21.3) & 16579 (21.3) & 15570 (20.0) & 15662 (20.2) & 77694 \\
\textbf{} & 90-94 & 5533 (16.6) & 6968 (20.8) & 7352 (22.0) & 6639 (19.9) & 6936 (20.7) & 33428 \\
\textbf{} & 95 plus & 1489 (16.0) & 1829 (19.6) & 2082 (22.3) & 1928 (20.7) & 2001 (21.4) & 9329 \\
\textbf{Total repeat medicines} & Mean (SD) & 6.6 (4.4) & 6 (4.2) & 5.4 (4.1) & 5 (3.9) & 4.6 (3.7) &  \\
\textbf{Medicines group} & 0-2 & 31564 (12.6) & 43305 (17.3) & 54615 (21.8) & 58546 (23.4) & 62339 (24.9) & 250369 \\
\textbf{} & 03-May & 40396 (15.5) & 51836 (19.8) & 56891 (21.8) & 55634 (21.3) & 56432 (21.6) & 261189 \\
\textbf{} & 06-Aug & 40686 (19.3) & 46814 (22.2) & 45050 (21.3) & 40573 (19.2) & 37997 (18.0) & 211120 \\
\textbf{} & 9+ & 49766 (24.7) & 49266 (24.4) & 41774 (20.7) & 33396 (16.6) & 27443 (13.6) & 201645 \\
\textbf{Receiving social care} &  & 24848 (23.1) & 27209 (25.3) & 22531 (21.0) & 18033 (16.8) & 14820 (13.8) & 107441 \\
\textbf{Receiving home care} &  & 10063 (24.4) & 10482 (25.5) & 8504 (20.7) & 6694 (16.3) & 5429 (13.2) & 41172 \\
\textbf{\textgreater{}=1 epsiode USC} &  & 59898 (21.0) & 63211 (22.2) & 59044 (20.7) & 52933 (18.6) & 49973 (17.5) & 285059 \\
\textbf{\textgreater{}= 1 unplanned hospital admission} &  & 33362 (22.9) & 33625 (23.1) & 30222 (20.7) & 25601 (17.6) & 22960 (15.8) & 145770 \\
\textbf{\textgreater{}= 1 A \& E attendance} &  & 41212 (22.2) & 42645 (23.0) & 37409 (20.1) & 32794 (17.7) & 31707 (17.1) & 185767 \\ \bottomrule
\end{tabular}%
\begin{tablenotes}
\item All values n(\%) unless otherwise stated 
\item \emph{Row}-wise percentages within grouped variables
\item USC = Unscheduled care
\end{tablenotes}
\end{threeparttable}%
}
\caption{Characteristics of study population by deprivation quintile 2013/14. N = 908,295}
\label{tab:results1-deprivation}
\end{table}

Breaking the cohort down by deprivation quintile reveals that there were fewer males and females in quintile one compared to other quintiles in 2015/16 (table \ref{tab:results1-deprivation}). There was, however, a higher percentage of deaths in this quintile than would be seen if deaths were distributed equally across the population - where a value of 20% would be expected. This pattern is also present for quintiles two and three. Only 15.7% of individuals resident in quintile 5 datazones died in 2015/16. 

There is little difference in average age across deprivation quintiles. Nevertheless, taking into account the lower overall numbers in quintile one, there are noticeabley lower percentages of individuals in the 90-94 and 95 and over age groups compared to other quintiles.
`
Individuals resident in quintile five datazones were more likely to receive less repeat medicines with lower mean values compared to other quintiles. A positive linear pattern is observable in the numbers and percentages receiving 0-2 repeat medicines from quintile 1 (31,564, 12.6%) to quintile 5 (62,339, 24.9%). The opposite (i.e. a negative linear trend) is observed for numbers and percentages of those receiving over 9 repeat medicines from quintile 1 (49,766, 24.7%) to quintile 5 (27,443, 13.6%). 

In terms of service use, the largest percentage of individuals receiving social care was found in quintile two with decreasing percentages to quintile five. The second highest percetnage of social care users were found in quintile one. 

\FloatBarrier
\subsection{Deprivation by age and sex}\label{subsec:dep-agesex}

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-1/06_age_sex_simd_plot.png}
    \caption{Breakdown of population by age group, sex, and SIMD decile (Decile 1 - most deprived)}
    \label{fig:results1-age_simd}
\end{figure}

Figure \ref{fig:results1-age_simd} shows that the percentage of the cohort living in SIMD decile 1 areas is below 10% for both sexes and decreases with age. Males in decile 2 also show lower percentages and a reduction in overall make-up of the cohort at older ages.  

\FloatBarrier
\subsection{Deprivation by local authority}\label{subsec:dep-la}

The percentage of individuals from the cohort living in each of the SIMD deciles by local authority area of residence in the financial year 2013/14 is shown in figure \ref{fig:results1-la-simd}. The figure reveals large variation in percentages in each decile according to each local authority. For example approximately one-third of individuals in the City of Edinburgh live in a decile 10 datazone (the most affluent) whilst a similar number live in a decile 1 datazone in the Glasgow City area. Fife and Renfrewshire councils (and to a lesser extent South Ayrshire and South Lanarkshire) show the most even distribution of population across deprivation deciles albeit with slight discrepancies. Island local authority areas do not contain datazones from all 10 deciles and tend to have populations concentrated in the middle of the spectrum (a pattern also present in the Highland council area with smaller percentages seen at the extremes).    

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-1/05_simd_la.png}
    \caption{Distribution of cohort by local authority and SIMD decile (Decile 1 - most deprived)}
    \label{fig:results1-la-simd}
\end{figure}

\FloatBarrier
\section{Social Care}\label{sec:res1-sc}

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-1/13_comb_3.png}
    \caption{Receipt of social care (a)}
    \label{fig:results1-sc-demos}
\end{figure}

Figure \ref{fig:results1-sc-demos} illustrates that age is a significant contributing factor to the receipt of any form of social care with the percentage of the population receiving care increasing rapidly from age 65 to 95 regardless of sex, deprivation status or repeat medicine group. Despite this, there are noticeable differences within each of these groups with females, those in more deprived SIMD deciles, and those with higher multimorbidity status (measured by medicine group) showing higher percentages of individuals receiving social care at all ages. Looking across groups reveals a higher percentage of individuals prescribed nine or more repeat medicines receive care than any other group at all but the oldest ages.   

The difference between the percentage of females and males receiving social care at all ages is evident in both SIMD deciles one and ten (figure \ref{fig:results1-sc-simd-split}) suggesting deprivation status does not account for the difference in sex. This is also true for medicine groups with increasing percentages seen for both deciles from low to high medicine groups.  

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-1/13_comb_4.png}
    \caption{Receipt of social care (b)}
    \label{fig:results1-sc-simd-split}
\end{figure}

\FloatBarrier
\section{Unplanned hospital admission}\label{sec:res1-adm}

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-1/13_comb_1.png}
    \caption{Unplanned hospital admission (a)}
    \label{fig:results1-adm-demos}
\end{figure}

Age is also an important driver in the percentage of individuals having at least one unplanned hospital admission (figure \ref{fig:results1-adm-demos}), although this is to a lesser extent than seen for social care. In contrast to social care receipt there is relatively little difference between sexes on unplanned admission. Similarly to social care, however, there are marked variations within deprivation and multimorbidty groups. Again, those in more deprived areas and with higher repeat medicine counts being more likely to have an unplanned admission at all ages.

The percentage of indiviudals in deciles one and ten with a hospital admission is broadly similar when split by sex (figure \ref{fig:results1-adm-simd-split}), although there is a marginally higher rate of men with an admission seen in ages 80 and under in decile 10. Multimorbidity has an effect across SIMD deciles with higher percentages of unplanned admission in both deciles one and ten as the number of repeat medicines increases. The gap between deciles one and ten also increses through low to high medicine groups. 

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-1/13_comb_2.png}
    \caption{Unplanned hospital admission (b)}
    \label{fig:results1-adm-simd-split}
\end{figure}
