---
output: pdf_document
---

\chapter{Results 2: Social Care and Multimorbidity}\label{ch:res2}
\thispagestyle{empty}

Introductory paragraph - Mention looking at any form of social care here (basically being in SCS), not home care or telecare specifically. Also reiterate the split in LAs by linkage rates (section \ref{stats-rq1}.)

\section{Overall Descriptive statistics}\label{subsec:res2-desc}

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-1/17_age_sex_scs_plot.png}
    \caption{Percentage of individuals receiving social care by sex at specific ages. 2013/14}
    \label{fig:res2-age-sex}
\end{figure}

Figure \ref{fig:res2-age-sex} shows that as age increases, higher percentages of indiviudals receive social care in financial year 2013/14. This pattern is consistent across all years of data. There is an uneven split by sex. Females have higher percentages of indivudals receiving any form of social care than males at all ages. At age 85, ~20% of males and ~32% of females received any form of social care. The difference between the percentage of males and females receiving care increases as age increases. There is mild reduction of this trend over the age of 90.   

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-1/18_age_simd_scs_plot.png}
    \caption{Percentage of indiviudals receiving social care by deprivation at specific ages. 2013/14}
    \label{fig:res2-simd}
\end{figure}

Figure \ref{fig:res2-simd} also shows the increase in the percentage of individuals receiving any form of social care as age increases. The split by SIMD decile shows that those resident in SIMD decile 1 datazones (most deprived) have higher percentages receiving social care at all ages (with the exceptions of ages 88 and 92). There is a clear gradient from decile 1 to decile 10 (most affluent) which has the lowest percentages at all ages (except at over the age of 90 where small numbers introduce wider variation).  

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-1/20_age_meds_scs_plot.png}
    \caption{Percentage of indiviudals receiving social care by multimorbidity (repeat medicine) group at specific ages. 2013/14}
    \label{fig:res2-meds-age}
\end{figure}

A third figure charting the increase in the percentage of individuals receiving any form of social care as age increases in 2013/14 is shown in figure \ref{fig:res2-meds-age}. The split in this figure shows that individuals receiving nine or more medicines have much higher percentages receiving social care at all ages. For example, ~5% of indivudals with 0-2 repeat medicines at age 80 receive some form of social care, whereas ~26% of individuals receiving 9 or more repeat medicines at age 80 are found in the SCS.


```{r models_1, eval=FALSE, echo=FALSE}
###################################################################################################
```

\FloatBarrier
\section{Models for councils with link rate greater than 96\%}\label{sec:96plus}

This section reports results of logistic regression models detailed in section \ref{stats-rq1}. As that section describes, the cohort was subset to include local authorities with similar levels of linkage to the population spine so that fair comparisons could be made across authorities. In this section nine local authorities are included in the logisitic regression models. Over 96% of the individuals in the SCS from these local authorites were matched to the population spine enabling linkage to other sources (table \ref{tab:linkage-rates}). 

In total, five models are reported - one for each financial year of data. As cohort data was subset to include indiviudals only from the nine local authorites with link rates over 96%, descriptive statistics of this subset are first reported. This is follwed by results of goodness of fit tests applied to each model of data. Model effects are reported as Average Partial Effects (APEs). As with previous results, financial year 2013/14 is used as a reference year where patterns are consistent across all years of data.

\subsection{Descriptive statistics}\label{subsec:res2-model1-desc}

Table \ref{tab:res2-chars1314} shows a breakdown of the cohort in financial year 2013/14. There were a total of 221,763 individuals from the nine included local authority areas in this financial year. A higher perecntage of females (15.6%) compared to males (8.5%) received social care. Only 3.1% of those aged 65-69 received social care whereas 71.2% of those over the age of 95 did so. SIMD was modelled as a continous variable. The median vale of 4 (IQR 4) was lower (more deprived) for those that received social care than for those that did not (Median 5, IQR 4). A small proportion (3.3%) of those in the lowest multimorbidity group (0-2 repeat medicines) received social care whereas over a quarter (26.4%) of those in the highest multimorbidity group received care. 

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{threeparttable}
\begin{tabular}{@{}llrrrr@{}}
\toprule
\textbf{Variable} & \textbf{Value} & \textbf{No Social Care} & \textbf{Social Care} & \textbf{Total} & \textbf{p-value} \\ \midrule
\textbf{Sex} & Male & 91792 (91.5) & 8538 (8.5) & 100330 & \textless{}0.001 \\
\textbf{} & Female & 102532 (84.4) & 18901 (15.6) & 121433 &  \\
\textbf{Age group} & 65-69 & 69238 (96.9) & 2205 (3.1) & 71443 & \textless{}0.001 \\
\textbf{} & 70-74 & 51808 (94.1) & 3222 (5.9) & 55030 &  \\
\textbf{} & 75-79 & 38603 (88.8) & 4854 (11.2) & 43457 &  \\
\textbf{} & 80-84 & 22838 (77.0) & 6822 (23.0) & 29660 &  \\
\textbf{} & 85-89 & 9114 (59.5) & 6216 (40.5) & 15330 &  \\
\textbf{} & 90-94 & 2399 (42.0) & 3318 (58.0) & 5717 &  \\
\textbf{} & 95 plus & 324 (28.8) & 802 (71.2) & 1126 &  \\
\textbf{\begin{tabular}[c]{@{}l@{}}SIMD \\ (continuous variable)\end{tabular}} & Median (IQR) & 5 (4) & 4 (4) &  & \textless{}0.001 \\
\textbf{Repeat medicines group} & 0-2 & 56488 (96.7) & 1918 (3.3) & 58406 & \textless{}0.001 \\
\textbf{} & 03-May & 58323 (91.9) & 5161 (8.1) & 63484 &  \\
\textbf{} & 06-Aug & 44023 (85.3) & 7614 (14.7) & 51637 &  \\
\textbf{} & over 9 & 35490 (73.6) & 12746 (26.4) & 48236 &  \\ \bottomrule
\end{tabular}%
\begin{tablenotes}
\item Total N used in model = 221,763
\item All values n(\%) unless otherwise stated 
\item \emph{Row}-wise percentages within grouped variables
\end{tablenotes}
\end{threeparttable}%
}
\caption{Characteristics of those receiving/not receiving social care for council with linkage rate 96\% - 99.9\%. 2013/14}
\label{tab:res2-chars1314}
\end{table}

As there was a change in the method of data collection from 2014/15 onwards (described in section \ref{subsec:source-sc}), percentages of indiviudals receiving social care in each local authority is reported across all years of data in table \ref{tab:res2-la}. With the exception of financial year 2011/12, Falkirk council has the highest proportion of individuals receiving care in all years. Whilst percentages of indiviudals receiving care are stable across years in some local authorities, there are noticeable increases in values for others from 2014/15 onwards (particularly in North Ayrshire, Dumfries & Galloway and Argyll & Bute). 

\begin{table}[h]
\centering
\begin{tabular}{@{}lrrrrr@{}}
\toprule
\textbf{Local Authority} & \textbf{2011/12} & \textbf{2012/13} & \textbf{2013/14} & \textbf{2014/15} & \textbf{2015/16} \\ \midrule
Falkirk & 14.8 & 16 & 15.8 & 16.8 & 16.5 \\
Angus & 15.8 & 14.7 & 14.1 & 14.6 & 12.8 \\
East Ayrshire & 13.7 & 13.9 & 14.1 & 14.8 & 14.8 \\
Inverclyde & 12.7 & 13 & 13 & 14.1 & 17.5 \\
South Lanarkshire & 12.2 & 11.9 & 12.9 & 13.7 & 13.2 \\
Stirling & 11.7 & 11.8 & 12.6 & 13.1 & 13.3 \\
North Ayrshire & 11.3 & 11.6 & 11.9 & 13.8 & 13.4 \\
Dumfries and Galloway & 9 & 8.7 & 8.1 & 10.6 & 11.3 \\
Argyll and Bute & 8.7 & 9.1 & 9.6 & 12.1 & 10.8 \\ \bottomrule
\end{tabular}
\caption{Percentage of over 65s receiving social care, by local authority}
\label{tab:res2-la}
\end{table}



\FloatBarrier
\subsection{Goodness of fit}\label{subsec:96plus-gof}

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrr@{}}
\toprule
\textbf{Financial Year} & \textbf{\begin{tabular}[c]{@{}r@{}}McFadden's \\ pseudo $R^{2}$\end{tabular}} & \textbf{Deviance} & \textbf{Log Liklihood} & \textbf{\begin{tabular}[c]{@{}r@{}}Deviance\\ (null model)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Log liklihood\\ (null model)\end{tabular}} \\ \midrule
\textbf{2011/12} & 0.252 & 116463.376 & -58231.688 & 155736.471 & -77868.236 \\
\textbf{2012/13} & 0.255 & 119558.387 & -59779.194 & 160465.022 & -80232.511 \\
\textbf{2013/14} & 0.256 & 123591.938 & -61795.969 & 166009.028 & -83004.514 \\
\textbf{2014/15} & 0.264 & 133386.433 & -66693.217 & 181158.55 & -90579.275 \\
\textbf{2015/16} & 0.268 & 133304.348 & -66652.174 & 182135.575 & -91067.788 \\ \bottomrule
\end{tabular}%
}
\caption{Goodness of fit for model applied to each year of data}
\label{tab:res2-model1-gof}
\end{table}

Table \ref{tab:res2-model1-gof} shows the results of diagnostic tests for the models applied to each financial year of data. McFadden's pseudo $R^2$ statistic ranges from 0.252 for the model applied to 2011/12 to 0.268 for the model applied to 2015/16. This suggests the the model has an excellent fit to all years of data and gets better in more recent years. 

\FloatBarrier
\subsection{Average Partial Effects}\label{subsec:96plus-AME}


\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_22_AME_plot.png}
    \caption{Average partial effect plot. Local authorites with linkage rate 96\% - 99.9\%}
    \label{fig:res2-model1-APE}
\end{figure}

Figure \ref{fig:res2-model1-APE} plots the Average Partial Effects (APEs) of each model. Age has the largest effect with the probability of receiving social care approximately 60% higher for those over 95 compared to those aged 65-69 in all models. Age bands 90-94, 85-89, and 80-84 also show very strong effects (~50%, ~30%, and ~17% higher probability of receiving social care). 

Multimorbidity has a lesser, but still strong, effect on recipt of social care. Those receiving 9 or more repeat medicines are ~17% more likely to receive social care compared to those receiving 0-2 repeat medicines. This is effect is a similar magnitude as the APE of being 80-84 compared to 65-69 years of age. Medicine groups 6-8 and 3-5 have resepctively ~7% and ~3% increased probability of receiving social care compared to those receiving 0-2 medicines. 

Tight CIs dificult to visualise


\begin{table}[h]
\centering
\begin{threeparttable}
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Variable} & \textbf{\begin{tabular}[c]{@{}r@{}}Average\\ Partial Effect\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Lower 95\%\\ confidence interval\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Upper 95\%\\ confidence interval\end{tabular}} \\ \midrule
Age: 95 Plus & 0.625 & 0.592 & 0.657 \\
Age: 90-94 & 0.477 & 0.463 & 0.49 \\
Age: 85-89 & 0.314 & 0.306 & 0.321 \\
Meds: over 9 & 0.163 & 0.159 & 0.167 \\
Age: 80-84 & 0.158 & 0.154 & 0.163 \\
Meds: 6-8 & 0.07 & 0.067 & 0.073 \\
Age: 75-79 & 0.061 & 0.058 & 0.064 \\
Female & 0.044 & 0.042 & 0.047 \\
Meds: 3-5 & 0.032 & 0.029 & 0.035 \\
Age: 70-74 & 0.022 & 0.019 & 0.024 \\
Angus & -0.002 & -0.008 & 0.005 \\
SIMD Decile: & -0.006 & -0.007 & -0.006 \\
East Ayrshire & -0.021 & -0.027 & -0.015 \\
Stirling & -0.022 & -0.029 & -0.015 \\
South Lanarkshire & -0.035 & -0.04 & -0.031 \\
North Ayrshire & -0.038 & -0.043 & -0.033 \\
Inverclyde & -0.045 & -0.051 & -0.039 \\
Argyll and Bute & -0.053 & -0.059 & -0.047 \\
Dumfries and Galloway & -0.071 & -0.076 & -0.066 \\ \bottomrule
\end{tabular}
\begin{tablenotes}
\item Reference group for age: 65-69
\item Reference group for meds: 0-2 repeat medicines
\item Reference group for sex: Male
\item Reference group for local authority: Falkirk
\end{tablenotes}
\end{threeparttable}
\caption{Average partial effects. 2013/14}
\label{tab:res2-model1-APE}
\end{table}

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_22_AME_time.png}
    \caption{Average partial effect across models (a)}
    \label{fig:res2-model1-APE-time}
\end{figure}

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_22_AMELA_time.png}
    \caption{Average partial effect across models (b)}
    \label{fig:res2-model1-APELA-time}
\end{figure}

```{r model_2, echo=FALSE, eval=FALSE}
###################################################################################################
```

\FloatBarrier
\section{Model for councils with link rate between 92\% and 95.9\%}\label{sec:9296}

\subsection{Descriptive statistics}\label{subsec:res2-model2-desc}

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{threeparttable}
\begin{tabular}{@{}llrrrr@{}}
\toprule
\textbf{Variable} & \textbf{Value} & \textbf{No Social Care} & \textbf{Social Care} & \textbf{Total} & \textbf{p-value} \\ \midrule
\textbf{Sex} & Male & 123403 (92.7) & 9767 (7.3) & 133170 & <0.001 \\
\textbf{} & Female & 149820 (87.2) & 22090 (12.8) & 171910 &  \\
\textbf{Age group} & 65-69 & 90061 (97.4) & 2439 (2.6) & 92500 & <0.001 \\
\textbf{} & 70-74 & 67124 (95.2) & 3371 (4.8) & 70495 &  \\
\textbf{} & 75-79 & 54207 (91.1) & 5292 (8.9) & 59499 &  \\
\textbf{} & 80-84 & 35853 (82.5) & 7604 (17.5) & 43457 &  \\
\textbf{} & 85-89 & 17993 (70.1) & 7684 (29.9) & 25677 &  \\
\textbf{} & 90-94 & 6603 (60.0) & 4400 (40.0) & 11003 &  \\
\textbf{} & 95 plus & 1382 (56.4) & 1067 (43.6) & 2449 &  \\
\textbf{\begin{tabular}[c]{@{}l@{}}SIMD\\ (continuous variable)\end{tabular}} & Median (IQR) & 6 (6) & 5 (6) &  & <0.001 \\
\textbf{Repeat medicines group} & 0-2 & 79000 (97.1) & 2339 (2.9) & 81339 & <0.001 \\
\textbf{} & 03-May & 79976 (92.7) & 6266 (7.3) & 86242 &  \\
\textbf{} & 06-Aug & 61183 (87.2) & 8980 (12.8) & 70163 &  \\
\textbf{} & over 9 & 53064 (78.8) & 14272 (21.2) & 67336 &  \\ \bottomrule
\end{tabular}%
\begin{tablenotes}
\item All values n(\%) unless otherwise stated 
\item \emph{Row}-wise percentages within grouped variables
\end{tablenotes}
\end{threeparttable}%
}
\caption{Characteristics of those receiving/not reveiving social care for council with linkage rate 92\% - 95.9\%. 2013/14}
\label{tab:res2-mod2-chars1314}
\end{table}

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrr@{}}
\toprule
\textbf{Local Authority} & \textbf{2011/12} & \textbf{2012/13} & \textbf{2013/14} & \textbf{2014/15} & \textbf{2015/16} \\ \midrule
Shetland Islands & 18.3 & 18.4 & 17.3 & 15.5 & 16.6 \\
Glasgow City & 14.3 & 14.7 & 11.9 & 11.4 & 11.8 \\
Na h-Eileanan Siar & 11.6 & 10.8 & 11.7 & 12.3 & 16.9 \\
East Renfrewshire & 10.2 & 9.4 & 11.3 & 13.8 & 18.6 \\
City of Edinburgh & 9.7 & 9.8 & 10.4 & 12.6 & 11.2 \\
East Dunbartonshire & 9.8 & 10.1 & 9.9 & 11.8 & 12.4 \\
Fife & 9.3 & 9.6 & 9.4 & 10.9 & 11.3 \\
Perth and Kinross & 8.5 & 9.4 & 8.7 & 13.9 & 11.3 \\
South Ayrshire & 8.5 & 8.8 & 8.8 & 10.5 & 12.1 \\ \bottomrule
\end{tabular}%
}
\caption{Percentage of over 65s receiving social care, by local authority}
\label{tab:res2-mod2-la}
\end{table}

\FloatBarrier

\subsection{Goodness of fit}\label{subsec:9296-gof}

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrr@{}}
\toprule
\textbf{Financial Year} & \textbf{\begin{tabular}[c]{@{}r@{}}McFadden's\\ pseudo $R^{2}$\end{tabular}} & \textbf{Deviance} & \textbf{Log Liklihood} & \textbf{\begin{tabular}[c]{@{}r@{}}Deviance\\ (null model)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Log Liklihood\\ (null model)\end{tabular}} \\ \midrule
\textbf{2011/12} & 0.195 & 161593.493 & -80796.747 & 200854.348 & -100427.174 \\
\textbf{2012/13} & 0.199 & 166393.132 & -83196.566 & 207622.681 & -103811.3405 \\
\textbf{2013/14} & 0.199 & 163548.526 & -81774.263 & 204215.311 & -102107.6555 \\
\textbf{2014/15} & 0.216 & 177081.266 & -88540.633 & 225815.462 & -112907.731 \\
\textbf{2015/16} & 0.219 & 178813.735 & -89406.867 & 228954.547 & -114477.2735 \\ \bottomrule
\end{tabular}%
}
\caption{Goodness of fit of model applied to each year of data}
\label{tab:res2-mod2-gof}
\end{table}

\FloatBarrier

\subsection{Average Partial Effects}\label{subsec:9296-AME}

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_23_AME_plot.png}
    \caption{Average partial effect plot. Local authorites with linkage rate 92\% - 95.9\%}
    \label{fig:res2-model2-APE}
\end{figure}

\begin{table}[h]
\centering
\begin{threeparttable}
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Variable} & \textbf{\begin{tabular}[c]{@{}r@{}}Average\\ Partial Effect\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Lower 95\%\\ confidence interval\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Upper 95\%\\ confidence interval\end{tabular}} \\ \midrule
Age: 95 Plus & 0.401 & 0.378 & 0.424 \\
Age: 90-94 & 0.33 & 0.32 & 0.34 \\
Age: 85-89 & 0.228 & 0.223 & 0.234 \\
Meds: over 9 & 0.127 & 0.123 & 0.13 \\
Age: 80-84 & 0.119 & 0.115 & 0.122 \\
Meds: 6-8 & 0.061 & 0.058 & 0.063 \\
Age: 75-79 & 0.049 & 0.046 & 0.051 \\
Female & 0.03 & 0.028 & 0.032 \\
Meds: 3-5 & 0.028 & 0.025 & 0.03 \\
Age: 70-74 & 0.017 & 0.015 & 0.019 \\
SIMD Decile: & -0.005 & -0.006 & -0.005 \\
East Renfrewshire & -0.075 & -0.092 & -0.058 \\
City of Edinburgh & -0.081 & -0.096 & -0.065 \\
Na h-Eileanan Siar & -0.083 & -0.1 & -0.066 \\
East Dunbartonshire & -0.085 & -0.102 & -0.069 \\
Perth and Kinross & -0.092 & -0.108 & -0.076 \\
Glasgow City & -0.097 & -0.112 & -0.081 \\
Fife & -0.099 & -0.115 & -0.084 \\
South Ayrshire & -0.104 & -0.12 & -0.088 \\ \bottomrule
\end{tabular}
\begin{tablenotes}
\item Reference group for age: 65-69
\item Reference group for meds: 0-2 repeat medicines
\item Reference group for sex: Male
\item Reference group for local authority: Shetland Islands
\end{tablenotes}
\end{threeparttable}
\caption{Average partial effects. 2013/14}
\label{tab:res2-mod2-APE}
\end{table}

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_23_AME_time.png}
    \caption{Average partial effect across models (a)}
    \label{fig:res2-model1-APE-time}
\end{figure}

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_23_AMELA_time.png}
    \caption{Average partial effect across models (b)}
    \label{fig:res2-model1-APELA-time}
\end{figure}

\section{Conclusion}\label{sec-res2-conclusion}
