---
output: pdf_document
---

\chapter{Results 2: Models}\label{ch:res2}
\thispagestyle{empty}

This chapter reports results of an analysis conducted to answer research questions 1 and 2 namely: 

In people over the age of 65 in Scotland:

1. 
\begin{enumerate}[noitemsep]
\item What are the socioeconomic, demographic, and geographic patterns in the use of social care? \item Is there an association between multimorbidity status and social care use? Does this vary by the patterns described in 1(a)?  
\end{enumerate}
2. 
\begin{enumerate}[noitemsep]
\item Is there an association in the use of social care services, multimorbidity status and unplanned admission to hospital?
\end{enumerate}

As described in section \ref{stats-rq1}, receipt of social care was measured by an individual having a record present in the SCS of any year. This could be in relation to any form of social care provided, or commissioned, by a local authority. As linkage rates from the SCS to the population spine varied at the local authority level, direct comparison of levels of care across all councils is impossible. For this reason two subsets, together representing ~58% of the overall cohort, are compared in relation to research question 1 in the first two sections of this chapter. The third and final section aims to answer research question 2. 

\FloatBarrier
\section{Social care - subset 1}

\subsection{Models for councils with link rate greater than 96\%}\label{sec:96plus}

In this section results of logistic regression models applied to the subset of individuals in local authorities where a match rate greater than 96% was achieved when linking the SCS to the population spine. Nine local authorities are included in the logistic regression models: Falkirk, Angus, East Ayrshire, Inverclyde, South Lanarkshire, Stirling, North Ayrshire, Dumfries & Galloway, and Argyll & Bute.   

In total, five models are reported - one for each financial year of data. As in the previous chapter, where results are broadly consistent across years, 2015/16 is used as a reference year to convey results. 

The R code for the model fitted to this first subset of data was as follows:

```{r sc_model, eval=FALSE, echo=TRUE}
glm(scs_flag ~           #Flag for social care as dependent variable
      sex*age_grp +      #Interaction : Sex and Age group
      simd*meds_grp +    #Interaction : SIMD and repeat medicines
      age_grp*meds_grp + #Interaction : Age group and repeat medicines
      council*simd,      #Interaction : Council and SIMD
    family = binomial(), #logistic regression
    data = thesis)       #identify which data to use
```

R automtically fits all included variables individually before adding interaction terms.  

\subsection{Descriptive statistics}\label{subsec:res2-model1-desc}

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{threeparttable}
\begin{tabular}{@{}llrrrr@{}}
\toprule
\textbf{Variable} & \textbf{Value} & \textbf{No Social Care} & \textbf{Social Care} & \textbf{Total} & \textbf{p-value} \\ \midrule
\textbf{Sex} & Male & 97151 (91.7) & 8834 (8.3) & 105985 & \textless{}0.001 \\
\textbf{} & Female & 109096 (85.0) & 19311 (15.0) & 128407 &  \\
\textbf{Age group} & 65-69 & 71604 (97.1) & 2129 (2.9) & 73733 & \textless{}0.001 \\
\textbf{} & 70-74 & 53870 (94.6) & 3052 (5.4) & 56922 &  \\
\textbf{} & 75-79 & 40083 (89.2) & 4867 (10.8) & 44950 &  \\
\textbf{} & 80-84 & 24936 (78.2) & 6970 (21.8) & 31906 &  \\
\textbf{} & 85-89 & 11302 (63.2) & 6593 (36.8) & 17895 &  \\
\textbf{} & 90-94 & 3579 (50.4) & 3523 (49.6) & 7102 &  \\
\textbf{} & 95 plus & 873 (46.3) & 1011 (53.7) & 1884 &  \\
\textbf{SIMD decile} & 1 - most deprived & 15102 (83.1) & 3066 (16.9) & 18168 & \textless{}0.001 \\
\textbf{} & 2 & 20121 (85.0) & 3543 (15.0) & 23664 &  \\
\textbf{} & 3 & 23024 (85.3) & 3974 (14.7) & 26998 &  \\
\textbf{} & 4 & 24301 (87.0) & 3621 (13.0) & 27922 &  \\
\textbf{} & 5 & 24673 (88.5) & 3196 (11.5) & 27869 &  \\
\textbf{} & 6 & 27751 (89.6) & 3234 (10.4) & 30985 &  \\
\textbf{} & 7 & 20881 (89.4) & 2470 (10.6) & 23351 &  \\
\textbf{} & 8 & 20264 (90.2) & 2195 (9.8) & 22459 &  \\
\textbf{} & 9 & 19711 (91.4) & 1853 (8.6) & 21564 &  \\
\textbf{} & 10 - most affluent & 10419 (91.3) & 993 (8.7) & 11412 &  \\
\textbf{Repeat medicines group} & 0-2 & 59467 (96.9) & 1898 (3.1) & 61365 & \textless{}0.001 \\
\textbf{} & 3-5 & 61424 (92.2) & 5189 (7.8) & 66613 &  \\
\textbf{} & 6-8 & 46699 (85.6) & 7852 (14.4) & 54551 &  \\
\textbf{} & 9+ & 38657 (74.5) & 13206 (25.5) & 51863 &  \\
\textbf{Local Authority} & Falkirk & 22839 (85.5) & 3858 (14.5) & 26697 & \textless{}0.001 \\
\textbf{} & Angus & 20814 (88.2) & 2793 (11.8) & 23607 &  \\
\textbf{} & East Ayrshire & 19294 (86.1) & 3122 (13.9) & 22416 &  \\
\textbf{} & Inverclyde & 12793 (84.9) & 2280 (15.1) & 15073 &  \\
\textbf{} & South Lanarkshire & 46724 (88.1) & 6304 (11.9) & 53028 &  \\
\textbf{} & Stirling & 13202 (87.7) & 1859 (12.3) & 15061 &  \\
\textbf{} & North Ayrshire & 23407 (88.3) & 3094 (11.7) & 26501 &  \\
\textbf{} & Dumfries and Galloway & 30568 (90.9) & 3073 (9.1) & 33641 &  \\
\textbf{} & Argyll and Bute & 16606 (90.4) & 1762 (9.6) & 18368 &  \\ \bottomrule
\end{tabular}%
\begin{tablenotes}
\item \footnotesize{Total N used in model = 234,392}
\item \footnotesize{All values n(\%) unless otherwise stated} 
\item \footnotesize{\emph{Row}-wise percentages within grouped variables}
\end{tablenotes}
\end{threeparttable}%
}
\caption{Subset 1. Characteristics of those receiving/not receiving social care. 2015/16}
\label{tab:res2-chars1516}
\end{table}

There were 234,392 individuals included included in the model for 2015/16 (table \ref{tab:res2-chars1516}) which accounts for 25.4% of the overall cohort in that year. There are similar numbers included in models for other years. Proportions of females/males, age groups, SIMD decile, and repeat medicine groups are broadly similar to the overall cohort as reported in chapter \ref{ch:results-1}. Also in concordance with results in chapter \ref{ch:results-1}, table \ref{tab:res2-chars1516} shows higher percentages of females received social care as well as patterns indicating increasing social care use through age groups (young to old), SIMD deciles (affluent to deprived), and repeat medicine groups (low to high). Statistically significant results are found for all variables with Chi-squared tests for differences within groups. 

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrr@{}}
\toprule
 & \multicolumn{3}{c}{\begin{tabular}[c]{@{}l@{}}Data collected during\\ census week only\end{tabular}} & \multicolumn{2}{c}{\begin{tabular}[c]{@{}l@{}}Expanded data\\ collection for \\ telecare and\\ alarm services\end{tabular}} \\ \cmidrule(lr){2-4}\cmidrule(lr){5-6}
 & \textbf{2011/12} & \textbf{2012/13} & \textbf{2013/14} & \textbf{2014/15} & \textbf{2015/16} \\ \midrule
\textbf{Falkirk} & 14.1 & 15 & 14.8 & 15.2 & 14.5 \\
\textbf{Angus} & 15.1 & 13.8 & 13.5 & 13.7 & 11.8 \\
\textbf{East Ayrshire} & 12.9 & 13.3 & 13.2 & 13.8 & 13.9 \\
\textbf{Inverclyde} & 12.1 & 12.5 & 12.4 & 13.1 & 15.1 \\
\textbf{South Lanarkshire} & 11.7 & 11.4 & 12.4 & 13 & 11.9 \\
\textbf{Stirling} & 10.9 & 11.1 & 11.8 & 12.2 & 12.3 \\
\textbf{North Ayrshire} & 10.8 & 11 & 11.1 & 11.4 & 11.7 \\
\textbf{Dumfries and Galloway} & 8.4 & 8 & 7.5 & 8.1 & 9.1 \\
\textbf{Argyll and Bute} & 8.2 & 8.7 & 9.2 & 9.6 & 9.6 \\ \bottomrule
\end{tabular}%
}
\caption{Subset 1. Percentage of over 65s receiving any form of social care}
\label{tab:res2-la}
\end{table}

There is notable variation in the percentage of individuals receiving social care depending on the local authority of residence ranging from 14.9% in Falkirk to 9.1% in Dumfries & Galloway. These values, unlike other variables in the models, can fluctuate over time as shown in table \ref{tab:res2-chars1516}. There is no discernable pattern across council areas, some show gradual decreases over time, others gradual increases, whilst others fluctuate in both directions. Generally, all councils saw an increase in the percentage of those receiving care from 2014/15 onwards reflecting the change in data collection to include individuals receiving community alarm or telecare services at any time in the financial year. However, this increase was not sustained in all areas: Falkirk, Angus, and South Lanarkshire showing reductions in 2015/16 compared to the previous year.  


\FloatBarrier
\subsection{Model results}\label{subsec:96plus-results}
 
\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_25_or_plot.pdf}
    \caption{Subset 1 - odds ratios}
    \label{fig:res2-subset1-OR}
\end{figure} 
 
Odds ratios for age groups and repeat medicine groups in the fitted model (for 2015/16) increase from young-to-old and low-to-high respectively (figure \ref{fig:res2-subset1-OR}). Females are shown to be more likely to receive social care than males, whilst the effect of SIMD modelled as a continuous variable is marginally negative which indicates an increase in SIMD decile from deprived to affluent is associated with reduced chance of receiving social care. All local authorities show reduced odds of receiving social care compared to the reference group (Falkirk), but not all differences are statistically significant. 

Given the difficulties in interpretation of interaction terms in logisitic regression models (discussed in section \ref{sec:methods-stats}), these are not shown in figure \ref{fig:res2-subset1-OR}. Tabulation of full model results for 2015/16 are available in table \ref{append-tab-2} in Appendix D. 

In order to aid interpetation of the magnitude of influece each variable has on the receipt of social care, Average Partial Effects (APEs) were calculated. These are discussed fully in section \ref{sec:methods-stats}. In short, APEs indicate the marginal effect of each value of a variable on the outcome taking into account all other variables and interactions. 

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_25_AME_plot.pdf}
    \caption{Subset 1. Average Partial Effects}
    \label{fig:res2-model1-APE}
\end{figure}

 
\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{threeparttable}
\begin{tabular}{@{}lrrrrr@{}}
\toprule
\textbf{Variable} & \textbf{\begin{tabular}[c]{@{}r@{}}Average Partial\\ Effect\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Standard\\ Error\end{tabular}} & \textbf{p-value} & \textbf{\begin{tabular}[c]{@{}r@{}}Lower 95\%\\ CI\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Upper 95\%\\ CI\end{tabular}} \\ \midrule
Age: 95 Plus & 0.466 & 0.014 & \textless{}0.001 & 0.439 & 0.493 \\
Age: 90-94 & 0.417 & 0.007 & \textless{}0.001 & 0.404 & 0.43 \\
Age: 85-89 & 0.283 & 0.004 & \textless{}0.001 & 0.276 & 0.291 \\
Meds: 9+ & 0.153 & 0.002 & \textless{}0.001 & 0.149 & 0.157 \\
Age: 80-84 & 0.151 & 0.002 & \textless{}0.001 & 0.147 & 0.155 \\
Meds: 6-8 & 0.067 & 0.002 & \textless{}0.001 & 0.063 & 0.07 \\
Age: 75-79 & 0.062 & 0.002 & \textless{}0.001 & 0.059 & 0.065 \\
Female & 0.04 & 0.001 & \textless{}0.001 & 0.038 & 0.043 \\
Meds: 3-5 & 0.028 & 0.002 & \textless{}0.001 & 0.025 & 0.031 \\
Age: 70-74 & 0.02 & 0.001 & \textless{}0.001 & 0.017 & 0.022 \\
SIMD Decile: & -0.006 & 0 & \textless{}0.001 & -0.006 & -0.005 \\
East Ayrshire & -0.01 & 0.003 & 0.001 & -0.015 & -0.004 \\
Stirling & -0.01 & 0.003 & 0.003 & -0.017 & -0.004 \\
Inverclyde & -0.012 & 0.003 & \textless{}0.001 & -0.018 & -0.006 \\
Angus & -0.016 & 0.003 & \textless{}0.001 & -0.022 & -0.01 \\
North Ayrshire & -0.029 & 0.003 & \textless{}0.001 & -0.034 & -0.024 \\
South Lanarkshire & -0.033 & 0.002 & \textless{}0.001 & -0.038 & -0.029 \\
Argyll and Bute & -0.041 & 0.003 & \textless{}0.001 & -0.047 & -0.035 \\
Dumfries and Galloway & -0.05 & 0.002 & \textless{}0.001 & -0.055 & -0.045 \\ \bottomrule
\end{tabular}
\begin{tablenotes}
\item \footnotesize{Reference group for age: 65-69}
\item \footnotesize{Reference group for meds: 0-2 repeat medicines}
\item \footnotesize{Reference group for local authority: Falkirk}
\end{tablenotes}
\end{threeparttable}%
}
\caption{Subset 1. Average partial effects. 2015/16}
\label{tab:res2-model1-APE}
\end{table}

APEs for each variable in each model (year) are plotted in figure \ref{fig:res2-model1-APE} with values for 2015/16 shown in table \ref{tab:res2-model1-APE}. Age has the largest positive effect of receiving social care with the probability approximately 45% higher for those over 95 compared to those aged 65-69 in all models (specifically 46.6% in 2015/16). Age bands 90-94, 85-89, and 80-84 also show very strong effects (~40%, ~28%, and ~15% higher probability of receiving social care across all models). 

Multimorbidity also has a strong positive effect on receipt of social care, although the magnitude of this effect is lower than that of higher age groups. Those receiving 9 or more repeat medicines have ~15% higher probability of receiving social care compared to those receiving 0-2 repeat medicines in all years (15.3% in 2015/16). This effect is a similar magnitude as the APE of being 80-84 compared to 65-69 years of age. Medicine groups 6-8 and 3-5 have respectively ~7% and ~3% increased probability of receiving social care compared to those receiving 0-2 medicines.

After controlling for other variables in the model females have an increased probability of receiving care of ~4% compared to males for all models (exactly 4% in 2015/16). SIMD decile was modelled as a continuous variable. The APE of -0.006 in 2015/16 (table \ref{tab:res2-model1-APE}) indicates the coefficient of the slope at the average value of SIMD (5.2, SD 2.6). The small value of this slope suggests SIMD decile had little effect on the receipt of social care. This can be seen in the conditional effect of SIMD shown in figure \ref{fig:res2-model1-simd} in Appendix D.

With the exception of higher age groups where numbers are small, local authorities display more variation in APEs across models than other variables over time. This is more easily visualised in figure \ref{fig:res2-model1-APE-time} which shows that the difference in APEs \emph{within} councils is fairly stable and shows some narrowing \emph{between} councils in 2015/16. Controlling for age, sex, deprivation, and multimorbidity attenuates some of the variations seen in table \ref{tab:res2-la} with the largest difference in probabilty across all nine included local authorities being 5% (comparing Dumfries & Galloway to the reference group: Falkirk).

\begin{figure}[]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_25_council_APE_plot.pdf}
    \caption{Subset 1. Local authority APE across models}
    \label{fig:res2-model1-APE-time}
\end{figure}


\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrrrrrr@{}}
\toprule
 & \multicolumn{1}{l}{} & \multicolumn{3}{c}{Fitted model} & \multicolumn{3}{c}{Null model} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\ \cmidrule(lr){3-5}\cmidrule(lr){6-8}
\textbf{\begin{tabular}[c]{@{}l@{}}Financial\\ Year\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}McFadden's\\ pseudo ${R^2}$\end{tabular}} & \textbf{Deviance} & \textbf{\begin{tabular}[c]{@{}r@{}}Log\\ Likelihood\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Degrees\\ of freedom\end{tabular}} & \textbf{Deviance} & \textbf{\begin{tabular}[c]{@{}r@{}}Log\\ Likelihood\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Degrees\\ of freedom\end{tabular}} & \textbf{AIC} & \textbf{BIC} \\ \midrule
\textbf{2011/12} & 0.218 & 122092.147 & -61046.073 & 218335 & 156192.796 & -78096.398 & 218389 & 122202.147 & 122768.319 \\
\textbf{2012/13} & 0.222 & 124626.619 & -62313.309 & 223999 & 160148.076 & -80074.038 & 224053 & 124736.619 & 125304.199 \\
\textbf{2013/14} & 0.222 & 128595.553 & -64297.777 & 228726 & 165344.874 & -82672.437 & 228780 & 128705.553 & 129274.282 \\
\textbf{2014/15} & 0.233 & 131849.309 & -65924.654 & 231616 & 171893.338 & -85946.669 & 231670 & 131959.309 & 132528.728 \\
\textbf{2015/16} & 0.236 & 131443.94 & -65721.97 & 234337 & 172080.12 & -86040.06 & 234391 & 131553.94 & 132124.002 \\ \bottomrule
\end{tabular}%
}
\caption{Subset 1. Goodness-of-fit}
\label{tab:res2-model1-gof}
\end{table}

Table \ref{tab:res2-model1-gof} shows the results of diagnostic tests for the models applied to each financial year of data. McFadden's pseudo $R^2$ statistic ranges from 0.218 for the model applied to 2011/12 to 0.236 for the model applied to 2015/16. As discussed in section \ref{stats-rq1}, this suggests the the model has an excellent fit to all years of data with improvement in more recent years.

```{r model_2, echo=FALSE, eval=FALSE}
###################################################################################################
```

\FloatBarrier
\section{Social Care - subset 2}

\subsection{Models for councils with link rate between 92\% and 95.9\%}\label{sec:9296}

In this section results of logistic regression models applied to the subset of individuals in local authorities where a match rate between 92% and 95.9% was achieved when linking the SCS to the population spine. Nine local authorities are included in the models: Shetland Islands, Glasgow City, Na h-Eilenanan Siar, East Renfrewshire, City of Edinburgh, East Dunbartonshire, Fife, Perth & Kinross, and South Ayrshire. 

In a similar fashion as the previous section, five models are reported - one for each financial year of data and where results are broadly consistent across years, 2015/16 is used as a reference year to convey results. Results here are restricted to show differences with the models fitted in the previous secion.  

The R code for the model fitted to the second subset was identical to that reported in the previous section, the underlying subset of data being the only change. 

\FloatBarrier
\subsection{Descriptive statistics}\label{subsec:res2-model2-desc}

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{threeparttable}
\begin{tabular}{@{}llrrrr@{}}
\toprule
\textbf{Variable} & \textbf{Value} & \textbf{No Social Care} & \textbf{Social Care} & \textbf{Total} & \textbf{p-value} \\ \midrule
\textbf{Sex} & Male & 127450 (92.8) & 9839 (7.2) & 137289 & \textless{}0.001 \\
\textbf{} & Female & 151392 (87.1) & 22323 (12.9) & 173715 &  \\
\textbf{Age group} & 65-69 & 92859 (97.3) & 2605 (2.7) & 95464 & \textless{}0.001 \\
\textbf{} & 70-74 & 68999 (95.3) & 3428 (4.7) & 72427 &  \\
\textbf{} & 75-79 & 54048 (91.5) & 5038 (8.5) & 59086 &  \\
\textbf{} & 80-84 & 36738 (83.0) & 7540 (17.0) & 44278 &  \\
\textbf{} & 85-89 & 18252 (70.4) & 7687 (29.6) & 25939 &  \\
\textbf{} & 90-94 & 6408 (58.7) & 4505 (41.3) & 10913 &  \\
\textbf{} & 95 plus & 1538 (53.1) & 1359 (46.9) & 2897 &  \\
\textbf{SIMD decile} & 1 & 31012 (86.8) & 4727 (13.2) & 35739 & \textless{}0.001 \\
\textbf{} & 2 & 25243 (87.6) & 3573 (12.4) & 28816 &  \\
\textbf{} & 3 & 21948 (86.5) & 3416 (13.5) & 25364 &  \\
\textbf{} & 4 & 23057 (87.7) & 3223 (12.3) & 26280 &  \\
\textbf{} & 5 & 25118 (88.7) & 3208 (11.3) & 28326 &  \\
\textbf{} & 6 & 21785 (89.4) & 2572 (10.6) & 24357 &  \\
\textbf{} & 7 & 23311 (90.9) & 2338 (9.1) & 25649 &  \\
\textbf{} & 8 & 26783 (91.3) & 2548 (8.7) & 29331 &  \\
\textbf{} & 9 & 34987 (92.1) & 3015 (7.9) & 38002 &  \\
\textbf{} & 10 & 45598 (92.8) & 3542 (7.2) & 49140 &  \\
\textbf{Repeat medicines group} & 0-2 & 82052 (97.3) & 2259 (2.7) & 84311 & \textless{}0.001 \\
\textbf{} & 3-5 & 81945 (93.1) & 6084 (6.9) & 88029 &  \\
\textbf{} & 6-8 & 61694 (87.2) & 9094 (12.8) & 70788 &  \\
\textbf{} & 9+ & 53151 (78.3) & 14725 (21.7) & 67876 &  \\
\textbf{Local Authority} & Shetland Islands & 3393 (86.1) & 548 (13.9) & 3941 & \textless{}0.001 \\
\textbf{} & Glasgow City & 67579 (88.4) & 8853 (11.6) & 76432 &  \\
\textbf{} & Na h-Eileanan Siar & 4716 (85.4) & 809 (14.6) & 5525 &  \\
\textbf{} & East Renfrewshire & 14084 (87.5) & 2005 (12.5) & 16089 &  \\
\textbf{} & City of Edinburgh & 62236 (91.0) & 6177 (9.0) & 68413 &  \\
\textbf{} & East Dunbartonshire & 18664 (92.0) & 1626 (8.0) & 20290 &  \\
\textbf{} & Fife & 59594 (89.0) & 7358 (11.0) & 66952 &  \\
\textbf{} & Perth and Kinross & 26295 (90.1) & 2893 (9.9) & 29188 &  \\
\textbf{} & South Ayrshire & 22281 (92.2) & 1893 (7.8) & 24174 &  \\ \bottomrule
\end{tabular}%
\begin{tablenotes}
\item \footnotesize{Total N used in model = 311,004}
\item \footnotesize{All values n(\%) unless otherwise stated} 
\item \footnotesize{\emph{Row}-wise percentages within grouped variables}
\end{tablenotes}
\end{threeparttable}%
}
\caption{Subset 2. Characteristics of those receiving/not receiving social care. 2015/16}
\label{tab:res2-2-chars1516}
\end{table}

In this second subset, the nine included councils represented approximately 33% of the overall cohort (e.g. 311,004 individuals in 2015/16, table \ref{tab:res2-2-chars1516}). There were smaller percentages of individuals receiving social care in all sociodemographic and multimorbidity groups compared to subset 1 (table \ref{tab:res2-chars1516}). For example, 7.2% of males and 12.9% of females received social care in subset 2 compared to 8.3% of males and 15% of females in subset 1. Less than 10% of individuals over 65 received social care in four of the nine councils in subset 2 including the second largest council in Scotland - City of Edinburgh. Only two councils in subset 1, Dumfries & Galloway and Argyll & Bute, provided less than 10% of over 65s with social care.  

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrr@{}}
\toprule
 & \multicolumn{3}{c}{\begin{tabular}[c]{@{}c@{}}Data collected during\\ census week only\end{tabular}} & \multicolumn{2}{c}{\begin{tabular}[c]{@{}c@{}}Expanded data\\ collection for\\ telecare and \\ alarm services\end{tabular}} \\ \cmidrule(lr){2-4}\cmidrule(lr){5-6}
 & \textbf{2011/12} & \textbf{2012/13} & \textbf{2013/14} & \textbf{2014/15} & \textbf{2015/16} \\ \midrule
\textbf{Shetland Islands} & 17.8 & 16.8 & 15.9 & 14.8 & 13.9 \\
\textbf{Glasgow City} & 14.2 & 14.7 & 11.9 & 11.3 & 11.6 \\
\textbf{Na h-Eileanan Siar} & 11.6 & 10.7 & 11.6 & 11.2 & 14.6 \\
\textbf{East Renfrewshire} & 10.2 & 8.7 & 11.2 & 10.4 & 12.5 \\
\textbf{City of Edinburgh} & 9.5 & 9.1 & 9.7 & 10.5 & 9 \\
\textbf{East Dunbartonshire} & 9.6 & 9.9 & 9.7 & 10.3 & 8 \\
\textbf{Fife} & 9 & 9.3 & 9.1 & 10.5 & 11 \\
\textbf{Perth and Kinross} & 8.1 & 8.4 & 7.6 & 9.3 & 9.9 \\
\textbf{South Ayrshire} & 8.1 & 8.4 & 8.5 & 7.8 & 7.8 \\ \bottomrule
\end{tabular}%
}
\caption{Subset 2. Percentage of over 65s receiving any form of social care}
\label{tab:res2-mod2-la}
\end{table}


Table \ref{tab:res2-mod2-la} reveals that Shetland Islands council provided a much higher percentage of over 65 with social care than other councils in most years although this rate reduced over time. Other council areas show fluctuating percentages of their populations receiving social care. The change in data collection from 2014/15 onwards resulted in some councils showing a higher rate receiving care, whilst there was a drop in the percetnage receiving care in other areas.

\FloatBarrier
\subsection{Model results}\label{subsec:9296-AME}

Odds ratios for indiviudal values show similar patterns, but lower values, to those found in subset 1. A full table of model results for subset 2 (2015/16) can be found in figure \ref{fig:res2-model2-or} and table \ref{append-tab-3} in Appendix D. 

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_24_AME_plot.pdf}
    \caption{Subset 2. Average partial effect plot}
    \label{fig:res2-model2-APE}
\end{figure}

Broadly similar patterns APEs are found for sociodemographic and multimorbidity group variables in subset 2 as were found in subset 1 (figure \ref{fig:res2-model2-APE}) although these are of lower magnitude throughout. The only slight exceptions were for the oldest two age groups and the highest multimorbidity group where APEs were slightly less and 85-89 age group where the APE was slightly higher compared to subset 1. 

\begin{table}[]
\centering
\resizebox{\textwidth}{!}{%
\begin{threeparttable}
\begin{tabular}{@{}lrrrrr@{}}
\toprule
\textbf{Variable} & \textbf{\begin{tabular}[c]{@{}r@{}}Average Partial\\ Effect\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Standard\\ Error\end{tabular}} & \textbf{p-value} & \textbf{\begin{tabular}[c]{@{}r@{}}Lower 95\%\\ CI\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Upper 95\%\\ CI\end{tabular}} \\ \midrule
Age: 95 Plus & 0.415 & 0.011 & 0 & 0.394 & 0.437 \\
Age: 90-94 & 0.333 & 0.005 & 0 & 0.323 & 0.343 \\
Age: 85-89 & 0.22 & 0.003 & 0 & 0.214 & 0.225 \\
Meds: 9+ & 0.133 & 0.002 & 0 & 0.129 & 0.136 \\
Age: 80-84 & 0.112 & 0.002 & 0 & 0.109 & 0.116 \\
Meds: 6-8 & 0.062 & 0.001 & 0 & 0.059 & 0.065 \\
Age: 75-79 & 0.044 & 0.001 & 0 & 0.041 & 0.046 \\
Female & 0.032 & 0.001 & 0 & 0.03 & 0.034 \\
Meds: 3-5 & 0.027 & 0.001 & 0 & 0.024 & 0.029 \\
Age: 70-74 & 0.015 & 0.001 & 0 & 0.013 & 0.018 \\
SIMD Decile: & -0.005 & 0 & 0 & -0.005 & -0.004 \\
Na h-Eileanan Siar & -0.014 & 0.008 & 0.074 & -0.03 & 0.001 \\
East Renfrewshire & -0.032 & 0.008 & 0 & -0.047 & -0.017 \\
Perth and Kinross & -0.042 & 0.007 & 0 & -0.056 & -0.028 \\
Fife & -0.044 & 0.007 & 0 & -0.058 & -0.031 \\
City of Edinburgh & -0.053 & 0.007 & 0 & -0.067 & -0.04 \\
Glasgow City & -0.06 & 0.007 & 0 & -0.074 & -0.047 \\
East Dunbartonshire & -0.072 & 0.007 & 0 & -0.086 & -0.057 \\
South Ayrshire & -0.076 & 0.007 & 0 & -0.089 & -0.062 \\ \bottomrule
\end{tabular}
\begin{tablenotes}
\item \footnotesize{Reference group for age: 65-69}
\item \footnotesize{Reference group for meds: 0-2 repeat medicines}
\item \footnotesize{Reference group for local authority: Shetland Islands}
\end{tablenotes}
\end{threeparttable}%
}
\caption{Subset 2. Average partial effects. 2015/16}
\label{tab:res2-model2-APE}
\end{table}

More variation is seen across years for council areas than was evident with subset 1 and noticeably larger confidence intervals can be seen in figure \ref{fig:res2-model2-APE}. 
The gradual increase in APEs reflects the large reduction in the percentage of individuals receiving care over time in the reference group (Shetland Islands) from a high starting point. This makes comparison to the reference group less informative. However, APEs can still be interpreted between other values (councils) in the variable. Excluding the referecnce group, there is little variation in APEs \emph(between) and \emph{within} councils, although there is some widening across councils in the most recent year of data. The difference in probability of receiving social care between Na h-Eileanan Siar and South Ayrshire is 6.2% in 2015/16. 

Table \ref{tab:res2-mod2-gof} shows the results of diagnostic tests for the models applied to each financial year of data. McFadden's pseudo $R^2$ statistic ranges from 0.196 for the model applied to 2011/12 to 0.209 for the model applied to 2015/16. The models applied to the most recent two years of data have an excellent fit, whereas the previous three models are only marginally below the threshold of 0.2 generally regarded as representing an excellent fit.

\begin{figure}[H]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_24_council_APE_plot.pdf}
    \caption{Subset 2. Local authority APE across models}
    \label{fig:res2-model2-APE-time}
\end{figure}


\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrrrrrr@{}}
\toprule
 & \multicolumn{1}{l}{} & \multicolumn{3}{c}{Fitted model} & \multicolumn{3}{c}{Null model} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\ \cmidrule(lr){3-5}\cmidrule(lr){6-8}
\textbf{\begin{tabular}[c]{@{}l@{}}Financial\\ Year\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}McFadden's\\ pseudo ${R^2}$\end{tabular}} & \textbf{Deviance} & \textbf{\begin{tabular}[c]{@{}r@{}}Log\\ Likelihood\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Degrees\\ of freedom\end{tabular}} & \textbf{Deviance} & \textbf{\begin{tabular}[c]{@{}r@{}}Log\\ Likelihood\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Degrees\\ of freedom\end{tabular}} & \textbf{AIC} & \textbf{BIC} \\ \midrule
\textbf{2011/12} & 0.196 & 159534.227 & -79767.114 & 293867 & 198430.468 & -99215.234 & 293921 & 159644.227 & 160226.736 \\
\textbf{2012/13} & 0.198 & 162331.401 & -81165.7 & 299514 & 202476.6 & -101238.3 & 299568 & 162441.401 & 163024.956 \\
\textbf{2013/14} & 0.199 & 159148.11 & -79574.055 & 305011 & 198718.545 & -99359.2725 & 305065 & 159258.11 & 159842.666 \\
\textbf{2014/15} & 0.204 & 164081.461 & -82040.731 & 308392 & 206217.188 & -103108.594 & 308446 & 164191.461 & 164776.623 \\
\textbf{2015/16} & 0.209 & 163509.29 & -81754.645 & 310949 & 206829.494 & -103414.747 & 311003 & 163619.29 & 164204.906 \\ \bottomrule
\end{tabular}%
}
\caption{Subset 2. Goodness-of-fit}
\label{tab:res2-mod2-gof}
\end{table}

\FloatBarrier
\section{Unplanned hospital admission}\label{sec:res3-usc}

In this section the dependent variable for modelling is any unplanned hospital admission in the six months following the end of each financial year. This time period was chosen to allow inclusion of social care (measured by presence in the SCS collected at the end of each financial year) as an independent variable in the model. 

As in previous sections in this chapter, logistic regression models are applied to each financial year of data, five models in total, and 2015/16 is used as a reference year where results are consistent across models. Exploratory models revealed Local Authority and Health Board area of residence had little impact on outcomes so they were omitted from the final models. This meant data was not subset as in previous sections and models were fitted to the whole cohort in each year. 

The R code for the model fitted to the data was as follows. 

```{r adm_model, eval=FALSE, echo=TRUE}
glm(adm_six ~            #Flag for unplanned hospital admission
      sex*age_grp +      #Interaction : Sex and Age group
      simd*meds_grp +    #Interaction : SIMD and repeat medicines
      age_grp*meds_grp + #Interaction : Age group and repeat medicines
      scs_flag*age_grp + #Interaction : Social care receipt and Age group
      usc_flag +         #Any USC use in previous financial year
      care_home,         #Flag indicating residence in a care home
    family = binomial(), #logistic regression
    data = thesis_usc)   #identify which data to use
```

\FloatBarrier
\subsection{Descriptive statistics}\label{subsec:res3-adm-desc}

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{threeparttable}
\begin{tabular}{@{}llrrrr@{}}
\toprule
\textbf{Variable} & \textbf{Value} & \textbf{\begin{tabular}[c]{@{}r@{}}No unplanned\\ admission\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}At least one\\ unplanned admission\end{tabular}} & \textbf{Total} & \textbf{p-value} \\ \midrule
\textbf{Sex} & Male & 362216 (90.9) & 36145 (9.1) & 398361 & 0.221 \\
\textbf{} & Female & 444060 (91.0) & 43910 (9.0) & 487970 &  \\
\textbf{Age group} & 65-69 & 264091 (94.8) & 14522 (5.2) & 278613 & \textless{}0.001 \\
\textbf{} & 70-74 & 196494 (93.2) & 14418 (6.8) & 210912 &  \\
\textbf{} & 75-79 & 152457 (90.4) & 16208 (9.6) & 168665 &  \\
\textbf{} & 80-84 & 106369 (86.9) & 15983 (13.1) & 122352 &  \\
\textbf{} & 85-89 & 58183 (83.1) & 11831 (16.9) & 70014 &  \\
\textbf{} & 90-94 & 22701 (80.2) & 5610 (19.8) & 28311 &  \\
\textbf{} & 95 plus & 5981 (80.1) & 1483 (19.9) & 7464 &  \\
\textbf{SIMD decile} & 1 - most deprived & 63364 (86.9) & 9527 (13.1) & 72891 & \textless{}0.001 \\
\textbf{} & 2 & 71876 (88.6) & 9247 (11.4) & 81123 &  \\
\textbf{} & 3 & 80768 (89.4) & 9602 (10.6) & 90370 &  \\
\textbf{} & 4 & 83216 (90.3) & 8947 (9.7) & 92163 &  \\
\textbf{} & 5 & 85840 (91.0) & 8456 (9.0) & 94296 &  \\
\textbf{} & 6 & 87908 (91.6) & 8027 (8.4) & 95935 &  \\
\textbf{} & 7 & 84892 (92.2) & 7215 (7.8) & 92107 &  \\
\textbf{} & 8 & 82592 (92.6) & 6602 (7.4) & 89194 &  \\
\textbf{} & 9 & 83433 (92.8) & 6519 (7.2) & 89952 &  \\
\textbf{} & 10 - most affluent & 82387 (93.3) & 5913 (6.7) & 88300 &  \\
\textbf{Repeat medicines group} & 0-2 & 233426 (96.4) & 8617 (3.6) & 242043 & \textless{}0.001 \\
\textbf{} & 3-5 & 236155 (93.7) & 15852 (6.3) & 252007 &  \\
\textbf{} & 6-8 & 181810 (89.8) & 20653 (10.2) & 202463 &  \\
\textbf{} & 9+ & 154885 (81.6) & 34933 (18.4) & 189818 &  \\
\textbf{Any form of social care} & No Social Care & 729724 (92.8) & 56233 (7.2) & 785957 & \textless{}0.001 \\
\textbf{} & Social Care & 76552 (76.3) & 23822 (23.7) & 100374 &  \\
\textbf{Any USC episode in previous year} & No USC & 601891 (94.5) & 34814 (5.5) & 636705 & \textless{}0.001 \\
\textbf{} & USC & 204385 (81.9) & 45241 (18.1) & 249626 &  \\
\textbf{Residence} & Not Care Home & 777019 (91.2) & 75348 (8.8) & 852367 & \textless{}0.001 \\
\textbf{} & Care Home & 23535 (83.7) & 4592 (16.3) & 28127 &  \\ \bottomrule
\end{tabular}%
\begin{tablenotes}
\item \footnotesize{Total N used in model = 886,331}
\item \footnotesize{All values n(\%) unless otherwise stated} 
\item \footnotesize{\emph{Row}-wise percentages within grouped variables}
\end{tablenotes}
\end{threeparttable}%
}
\caption{Characteristics of those with/without any unplanned hospital admission in the six months following end of 2015/16}
\label{tab:res3adm-chars1516}
\end{table}

In 2015/16, 886,331 individuals were included in the model representing 95.9% of the cohort in that year (table \ref{tab:res3adm-chars1516}) with only those that died during the financial year being omitted from models. As such the sociodemographic make up is representative of the overall cohort. Chi-squared testing reveals no statistically significant difference between the numbers of females and males with an unplanned admission. There are however, statistically significant diffrences in all other groups with older, more deprived, and individuals with higher multimorbidity being more likely to have an admission. Those in receipt of social care, with an episode of USC in the previous year, and those resident in a care home were also more likely to have an unplanned hospital admission.   


\FloatBarrier
\subsection{Model results}

Figure \ref{fig:res3-adm-OR} and table \ref{app-tab4} in Appendix D show odds ratios for individual variables in model fitted to data for 2015/16. 

The APEs for models with unplanned admission to hospital as dependent variable are much lower in magnitude for all variables (figure \ref{fig:res3-adm-APE} and table \ref{tab:res3-adm-ape1516}) compared with models fitted in sections \ref{subsec:96plus-results} and \ref{subsec:9296-AME} where social care was the dependent variable. 

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-3/adm_model_3_AME_plot.pdf}
    \caption{Unplanned hospital admissions. Average Partial Effects}
    \label{fig:res3-adm-APE}
\end{figure}

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{threeparttable}
\begin{tabular}{@{}lrrrrr@{}}
\toprule
\textbf{Variable} & \textbf{\begin{tabular}[c]{@{}r@{}}Average\\ Partial Effect\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Standard\\ Error\end{tabular}} & \textbf{p-value} & \textbf{\begin{tabular}[c]{@{}r@{}}Lower 95\%\\ CI\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Upper 95\%\\ CI\end{tabular}} \\ \midrule
\textbf{Meds: 9+} & 0.083 & 0.001 & \textless{}0.001 & 0.081 & 0.085 \\
\textbf{Age: 90-94} & 0.075 & 0.003 & \textless{}0.001 & 0.07 & 0.08 \\
\textbf{USC} & 0.073 & 0.001 & \textless{}0.001 & 0.072 & 0.075 \\
\textbf{Social Care} & 0.065 & 0.001 & \textless{}0.001 & 0.063 & 0.067 \\
\textbf{Age: 95 Plus} & 0.06 & 0.005 & \textless{}0.001 & 0.051 & 0.07 \\
\textbf{Age: 85-89} & 0.057 & 0.001 & \textless{}0.001 & 0.054 & 0.06 \\
\textbf{Age: 80-84} & 0.037 & 0.001 & \textless{}0.001 & 0.034 & 0.039 \\
\textbf{Meds: 6-8} & 0.037 & 0.001 & \textless{}0.001 & 0.035 & 0.039 \\
\textbf{Age: 75-79} & 0.021 & 0.001 & \textless{}0.001 & 0.019 & 0.023 \\
\textbf{Meds: 3-5} & 0.015 & 0.001 & \textless{}0.001 & 0.014 & 0.017 \\
\textbf{Age: 70-74} & 0.007 & 0.001 & \textless{}0.001 & 0.005 & 0.009 \\
\textbf{SIMD Decile:} & -0.003 & 0 & \textless{}0.001 & -0.003 & -0.003 \\
\textbf{Care Home} & -0.008 & 0.001 & \textless{}0.001 & -0.011 & -0.006 \\
\textbf{Female} & -0.016 & 0.001 & \textless{}0.001 & -0.017 & -0.015 \\ \bottomrule
\end{tabular}%
\begin{tablenotes}
\item \footnotesize{Reference group for age: 65-69}
\item \footnotesize{Reference group for meds: 0-2 repeat medicines}
\end{tablenotes}
\end{threeparttable}%
}
\caption{Unplanned hospital admission. Average Partial Effects. 2015/16}
\label{tab:res3-adm-ape1516}
\end{table}

High multimorbididty, measured by prescription of 9 or more repeat medicines, has the largest magnitude of APE in the unplanned hospital admission model in all years (an 8.3% increased probability of having an unplanned admission compared to those with 0-2 repeat medicines in 2015/16, table \ref{tab:res3-adm-ape1516}). Other multimorbidity groups also show increases in probability of admission compared to the reference value with a linear increase from low-to-high multimorbidity. 

In terms of other service use, having any USC episode in the previous financial year is associated with a ~7.5% increase in probability compared to those with no previous USC use. Similarly, receipt of social care measured in the SCS is associated with a ~6.5% increase in the probability of having an unplanned hospital admission compared to those with no social care. There is a split in the direction of the effect of living in a care home after 2013/14. After this year being in a care home is associated with a decrease in the probability of unplanned admission of almost 1% whilst in previous years the probability had increased by almost 1%. 

Higher age groups are also associated with higher APEs compared to other variables in the models though there are wider confidence intervals than seen with social care models. Furthermore, the magnitude of the effect of increasing age is much lower than the effect on receipt of social care. For example, in 2015/16 being 90-94 is associated with a 7.5% increase probability of having an unplanned hospital admission compared to being 65-69. The corresponding value for receiving social care in that age group is a 41.7% increased probability (subset 1, section \ref{subsec:96plus-results}). Otherwise, linear increases in APEs are seen through age groups.   

Despite no statistically significant difference in the number of males and females with an unplanned admission in descriptive table \ref{tab:res3adm-chars1516}, modelling reveals that females are almost 2% less likely to have an admission after adjusting for other variables (1.6% in 2015/16). 

SIMD decile, again fitted as a continous variable, has a very small effect on hospital admission after accounting for other variables. The coefficient of the slope at the average value of SIMD (5.6, SD 2.8) is -0.003 (table \ref{tab:res3-adm-ape1516}) indicating a very flat line which can be viewed in figure \ref{fig:res3-adm-simd} in Appendix D

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrrrrrr@{}}
\toprule
 & \multicolumn{1}{l}{} & \multicolumn{3}{c}{Fitted model} & \multicolumn{3}{c}{Null model} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\ \cmidrule(lr){3-5}\cmidrule(lr){6-8}
\textbf{\begin{tabular}[c]{@{}l@{}}Financial\\ Year\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}McFadden's\\ pseudo ${R^2}$\end{tabular}} & \textbf{Deviance} & \textbf{\begin{tabular}[c]{@{}r@{}}Log\\ Likelihood\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Degrees\\ of freedom\end{tabular}} & \textbf{Deviance} & \textbf{\begin{tabular}[c]{@{}r@{}}Log\\ Likelihood\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Degrees\\ of freedome\end{tabular}} & \textbf{AIC} & \textbf{BIC} \\ \midrule
\textbf{2011/12} & 0.113 & 447862.4 & -223931.2 & 821187.0 & 504865.0 & -252432.5 & 821234.0 & 447958.4 & 448516.1 \\
\textbf{2012/13} & 0.112 & 454527.8 & -227263.9 & 841129.0 & 512028.3 & -256014.2 & 841176.0 & 454623.8 & 455182.7 \\
\textbf{2013/14} & 0.114 & 466492.6 & -233246.3 & 858910.0 & 526463.8 & -263231.9 & 858957.0 & 466588.6 & 467148.4 \\
\textbf{2014/15} & 0.111 & 471115.4 & -235557.7 & 870332.0 & 530199.5 & -265099.8 & 870379.0 & 471211.4 & 471771.8 \\
\textbf{2015/16} & 0.113 & 475554.4 & -237777.2 & 880446.0 & 535977.2 & -267988.6 & 880493.0 & 475650.4 & 476211.4 \\ \bottomrule
\end{tabular}%
}
\caption{Unplanned hospital admission. Goodness-of-fit}
\label{res3-adm-gof}
\end{table}

Table \ref{res3-adm-gof} shows the results of diagnostic tests for each model. The value of McFadden's $R^2$ is fairly stable over time varying from 0.111 to 0.114. These values are well below those seen for social care models and represent a poorer fit to the data in comparison. 

