---
output: pdf_document
---

\chapter{Results 2: Models}\label{ch:res2}
\thispagestyle{empty}

This chapter reports results of an analysis conducted to answer research question 1 (link to section). As described in section \ref{stats-rq1}, receipt of social care was measured by an individual having a record present in the SCS of any year. This could be in relation to any form of social care provided, or commissioned, by a local authority. 

The chapter is split into three sections. The first reports descriptive statistics of receipt of social care for the whole cohort using financial year 2013/14 as a reference year. Plots showing the relationship between the receipt of social care and age, sex, deprivation status, and multimorbidity status are provided. 

The remaining sections report results of logistic regression models applied to each year of data for two subsets of the main cohort. The models in the second section were run on a subset of individuals from the nine local authorities where a match rate of over 96% was achieved when linking the SCS to the population spine. The models in the third section were applied to the subset of individuals from the nine local authorities with a match rate of 92-95.9%. These subsets were used to enable comparison of local authorities with similar match rates (thus, similar levels of missing data - section \ref{stats-rq1}) 

\FloatBarrier
\section[Section Title, Section subtitle]{Social care - subset 1\\ {\large Models for councils with link rate greater than 96\%}}\label{sec:96plus}

In this section results of logistic regression models applied to the subset of individuals in local authorities where a match rate greater than 96% was achieved when linking the SCS to the population spine. Nine local authorities are included in the logistic regression models: Falkirk, Angus, East Ayrshire, Inverclyde, South Lanarkshire, Stirling, North Ayrshire, Dumfries & Galloway, and Argyll & Bute.   

In total, five models are reported - one for each financial year of data. As in the previous chapter, where results are broadly consistent across years, 2015/16 is used as a reference year to convey results. 

The R code for the model fitted to this first subset of data was as follows:

```{r sc_model, eval=FALSE, echo=TRUE}
glm(scs_flag ~           #Flag for social care as dependent variable
      sex*age_grp +      #Interaction : Sex and Age group
      simd*meds_grp +    #Interaction : SIMD and repeat medicines
      age_grp*meds_grp + #Interaction : Age group and repeat medicines
      council*simd,      #Interaction : Council and SIMD
    family = binomial(), #logistic regression
    data = thesis)       #identify which data to use
```

R automtically fits all included variables individually before adding interaction terms.  

\subsection{Descriptive statistics}\label{subsec:res2-model1-desc}

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{threeparttable}
\begin{tabular}{@{}llrrrr@{}}
\toprule
\textbf{Variable} & \textbf{Value} & \textbf{No Social Care} & \textbf{Social Care} & \textbf{Total} & \textbf{p-value} \\ \midrule
\textbf{Sex} & Male & 97151 (91.7) & 8834 (8.3) & 105985 & \textless{}0.001 \\
\textbf{} & Female & 109096 (85.0) & 19311 (15.0) & 128407 &  \\
\textbf{Age group} & 65-69 & 71604 (97.1) & 2129 (2.9) & 73733 & \textless{}0.001 \\
\textbf{} & 70-74 & 53870 (94.6) & 3052 (5.4) & 56922 &  \\
\textbf{} & 75-79 & 40083 (89.2) & 4867 (10.8) & 44950 &  \\
\textbf{} & 80-84 & 24936 (78.2) & 6970 (21.8) & 31906 &  \\
\textbf{} & 85-89 & 11302 (63.2) & 6593 (36.8) & 17895 &  \\
\textbf{} & 90-94 & 3579 (50.4) & 3523 (49.6) & 7102 &  \\
\textbf{} & 95 plus & 873 (46.3) & 1011 (53.7) & 1884 &  \\
\textbf{SIMD decile} & 1 - most deprived & 15102 (83.1) & 3066 (16.9) & 18168 & \textless{}0.001 \\
\textbf{} & 2 & 20121 (85.0) & 3543 (15.0) & 23664 &  \\
\textbf{} & 3 & 23024 (85.3) & 3974 (14.7) & 26998 &  \\
\textbf{} & 4 & 24301 (87.0) & 3621 (13.0) & 27922 &  \\
\textbf{} & 5 & 24673 (88.5) & 3196 (11.5) & 27869 &  \\
\textbf{} & 6 & 27751 (89.6) & 3234 (10.4) & 30985 &  \\
\textbf{} & 7 & 20881 (89.4) & 2470 (10.6) & 23351 &  \\
\textbf{} & 8 & 20264 (90.2) & 2195 (9.8) & 22459 &  \\
\textbf{} & 9 & 19711 (91.4) & 1853 (8.6) & 21564 &  \\
\textbf{} & 10 - most affluent & 10419 (91.3) & 993 (8.7) & 11412 &  \\
\textbf{Repeat medicines group} & 0-2 & 59467 (96.9) & 1898 (3.1) & 61365 & \textless{}0.001 \\
\textbf{} & 3-5 & 61424 (92.2) & 5189 (7.8) & 66613 &  \\
\textbf{} & 6-8 & 46699 (85.6) & 7852 (14.4) & 54551 &  \\
\textbf{} & 9+ & 38657 (74.5) & 13206 (25.5) & 51863 &  \\
\textbf{Local Authority} & Falkirk & 22839 (85.5) & 3858 (14.5) & 26697 & \textless{}0.001 \\
\textbf{} & Angus & 20814 (88.2) & 2793 (11.8) & 23607 &  \\
\textbf{} & East Ayrshire & 19294 (86.1) & 3122 (13.9) & 22416 &  \\
\textbf{} & Inverclyde & 12793 (84.9) & 2280 (15.1) & 15073 &  \\
\textbf{} & South Lanarkshire & 46724 (88.1) & 6304 (11.9) & 53028 &  \\
\textbf{} & Stirling & 13202 (87.7) & 1859 (12.3) & 15061 &  \\
\textbf{} & North Ayrshire & 23407 (88.3) & 3094 (11.7) & 26501 &  \\
\textbf{} & Dumfries and Galloway & 30568 (90.9) & 3073 (9.1) & 33641 &  \\
\textbf{} & Argyll and Bute & 16606 (90.4) & 1762 (9.6) & 18368 &  \\ \bottomrule
\end{tabular}%
\begin{tablenotes}
\item \footnotesize{Total N used in model = 234,392}
\item \footnotesize{All values n(\%) unless otherwise stated} 
\item \footnotesize{\emph{Row}-wise percentages within grouped variables}
\end{tablenotes}
\end{threeparttable}%
}
\caption{Subset 1. Characteristics of those receiving/not receiving social care. 2015/16}
\label{tab:res2-chars1516}
\end{table}

There were 234,392 individuals included included in the model for 2015/16 (table \ref{tab:res2-chars1516}) which accounts for 25.4% of the overall cohort in that year. There are similar numbers included in models for other years. Proportions of females/males, age groups, SIMD decile, and repeat medicine groups are broadly similar to the overall cohort as reported in chapter \ref{ch:results-1}. Also in concordance with results in chapter \ref{ch:results-1}, table \ref{tab:res2-chars1516} shows higher percentages of females received social care as well as patterns indicating increasing social care use through age groups (young to old), SIMD deciles (affluent to deprived), and repeat medicine groups (low to high). Statistically significant results are found for all variables with Chi-squared tests for differences within groups. 

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrr@{}}
\toprule
 & \multicolumn{3}{c}{\begin{tabular}[c]{@{}l@{}}Data collected during\\ census week only\end{tabular}} & \multicolumn{2}{c}{\begin{tabular}[c]{@{}l@{}}Expanded data\\ collection for \\ telecare and\\ alarm services\end{tabular}} \\ \cmidrule(lr){2-4}\cmidrule(lr){5-6}
 & \textbf{2011/12} & \textbf{2012/13} & \textbf{2013/14} & \textbf{2014/15} & \textbf{2015/16} \\ \midrule
\textbf{Falkirk} & 14.1 & 15 & 14.8 & 15.2 & 14.5 \\
\textbf{Angus} & 15.1 & 13.8 & 13.5 & 13.7 & 11.8 \\
\textbf{East Ayrshire} & 12.9 & 13.3 & 13.2 & 13.8 & 13.9 \\
\textbf{Inverclyde} & 12.1 & 12.5 & 12.4 & 13.1 & 15.1 \\
\textbf{South Lanarkshire} & 11.7 & 11.4 & 12.4 & 13 & 11.9 \\
\textbf{Stirling} & 10.9 & 11.1 & 11.8 & 12.2 & 12.3 \\
\textbf{North Ayrshire} & 10.8 & 11 & 11.1 & 11.4 & 11.7 \\
\textbf{Dumfries and Galloway} & 8.4 & 8 & 7.5 & 8.1 & 9.1 \\
\textbf{Argyll and Bute} & 8.2 & 8.7 & 9.2 & 9.6 & 9.6 \\ \bottomrule
\end{tabular}%
}
\caption{Subset 1. Percentage of over 65s receiving any form of social care}
\label{tab:res2-la}
\end{table}

There is notable variation in the percentage of individuals receiving social care depending on the local authority of residence ranging from 14.9% in Falkirk to 9.1% in Dumfries & Galloway. These values, unlike other variables in the models, can fluctuate over time as shown in table \ref{tab:res2-chars1516}. There is no discernable pattern across council areas, some show gradual decreases over time, others gradual increases, whilst others fluctuate in both directions. Generally, all councils saw an increase in the percentage of those receiving care from 2014/15 onwards reflecting the change in data collection to include individuals receiving community alarm or telecare services at any time in the financial year. However, this increase was not sustained in all areas: Falkirk, Angus, and South Lanarkshire showing reductions in 2015/16 compared to the previous year.  


\FloatBarrier
\subsection{Model results}\label{subsec:96plus-results}
 
\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_25_or_plot.pdf}
    \caption{Subset 1 - odds ratios}
    \label{fig:res2-subset1-OR}
\end{figure} 
 
Odds ratios for age groups and repeat medicine groups in the fitted model increase from young-to-old and low-to-high respectively (figure \ref{fig:res2-subset1-OR}). Females are shown to be more likely to receive social care than males, whilst the effect of SIMD modelled as a continuous variable is marginally negative which indicates an increase in SIMD decile from deprived to affluent is associated with reduced chance of receiving social care. All local authorities show reduced odds of receiving social care compared to the reference group (Falkirk), but not differences are statistically significant. 

Given the difficulties in interpretation of interaction terms in logisitic regression models (discussed in section \ref{sec:methods-stats}), these are not shown in figure \ref{fig:res2-subset1-OR}. Tabulation of full model results are available in table \ref{append-tab-2} in Appendix D. 

In order to aid interpetation of the magnitude of influece each variable has on the receipt of social care, Average Partial Effects (APEs) were calculated. These are discussed fully in section \ref{sec:methods-stats}. In short, APEs indicate the marginal effect of each value of a variable on the outcome taking into account all other variables and interactions. 

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_25_AME_plot.pdf}
    \caption{Subset 1. Average Partial Effects}
    \label{fig:res2-model1-APE}
\end{figure}

 
\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{threeparttable}
\begin{tabular}{@{}lrrrrr@{}}
\toprule
\textbf{Variable} & \textbf{\begin{tabular}[c]{@{}r@{}}Average Partial\\ Effect\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Standard\\ Error\end{tabular}} & \textbf{p-value} & \textbf{\begin{tabular}[c]{@{}r@{}}Lower 95\%\\ CI\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Upper 95\%\\ CI\end{tabular}} \\ \midrule
Age: 95 Plus & 0.466 & 0.014 & \textless{}0.001 & 0.439 & 0.493 \\
Age: 90-94 & 0.417 & 0.007 & \textless{}0.001 & 0.404 & 0.43 \\
Age: 85-89 & 0.283 & 0.004 & \textless{}0.001 & 0.276 & 0.291 \\
Meds: 9+ & 0.153 & 0.002 & \textless{}0.001 & 0.149 & 0.157 \\
Age: 80-84 & 0.151 & 0.002 & \textless{}0.001 & 0.147 & 0.155 \\
Meds: 6-8 & 0.067 & 0.002 & \textless{}0.001 & 0.063 & 0.07 \\
Age: 75-79 & 0.062 & 0.002 & \textless{}0.001 & 0.059 & 0.065 \\
Female & 0.04 & 0.001 & \textless{}0.001 & 0.038 & 0.043 \\
Meds: 3-5 & 0.028 & 0.002 & \textless{}0.001 & 0.025 & 0.031 \\
Age: 70-74 & 0.02 & 0.001 & \textless{}0.001 & 0.017 & 0.022 \\
SIMD Decile: & -0.006 & 0 & \textless{}0.001 & -0.006 & -0.005 \\
East Ayrshire & -0.01 & 0.003 & 0.001 & -0.015 & -0.004 \\
Stirling & -0.01 & 0.003 & 0.003 & -0.017 & -0.004 \\
Inverclyde & -0.012 & 0.003 & \textless{}0.001 & -0.018 & -0.006 \\
Angus & -0.016 & 0.003 & \textless{}0.001 & -0.022 & -0.01 \\
North Ayrshire & -0.029 & 0.003 & \textless{}0.001 & -0.034 & -0.024 \\
South Lanarkshire & -0.033 & 0.002 & \textless{}0.001 & -0.038 & -0.029 \\
Argyll and Bute & -0.041 & 0.003 & \textless{}0.001 & -0.047 & -0.035 \\
Dumfries and Galloway & -0.05 & 0.002 & \textless{}0.001 & -0.055 & -0.045 \\ \bottomrule
\end{tabular}
\begin{tablenotes}
\item \footnotesize{Reference group for age: 65-69}
\item \footnotesize{Reference group for meds: 0-2 repeat medicines}
\item \footnotesize{Reference group for local authority: Falkirk}
\end{tablenotes}
\end{threeparttable}%
}
\caption{Subset 1. Average partial effects. 2015/16}
\label{tab:res2-model1-APE}
\end{table}

APEs for each variable in each model (year) are plotted in figure \ref{fig:res2-model1-APE} with values for 2015/16 shown in table \ref{tab:res2-model1-APE}. Age has the largest positive effect of receiving social care with the probability approximately 45% higher for those over 95 compared to those aged 65-69 in all models (specifically 46.6% in 2015/16). Age bands 90-94, 85-89, and 80-84 also show very strong effects (~40%, ~28%, and ~15% higher probability of receiving social care across all models). 

Multimorbidity also has a strong positive effect on receipt of social care, although the magnitude if this effect is lower than that of higher age groups. Those receiving 9 or more repeat medicines are ~15% more likely to receive social care compared to those receiving 0-2 repeat medicines in all years (15.3% in 2015/16). This effect is a similar magnitude as the APE of being 80-84 compared to 65-69 years of age. Medicine groups 6-8 and 3-5 have respectively ~7% and ~3% increased probability of receiving social care compared to those receiving 0-2 medicines.

After controlling for other variables in the model females have an increased probability of receiving care of ~4% compared to males for all models (exactly 4% in 2015/16).

After adjusting for all other variables in the model there remains variation in the likelihood of receiving social care across local authorities. In some cases this variation is small (e.g. in 2015/16 there is less than 2% decrease in the probability of receiving social care in East Ayrshire, Stirling, Inverclyde, and Angus compared to Falkirk) but can also have larger values (e.g. a 5% decrease in probability of receiving social care in Dumfries & Galloway compared to Falkirk). 

With the exception of higher age groups where numbers are small, local authorities display more variation in APEs across models than other variables. This is more easily visualised in figure \ref{fig:res2-model1-APE-time} 

\begin{figure}[]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_25_council_APE_plot.pdf}
    \caption{Subset 1. Local authority APE across models}
    \label{fig:res2-model1-APE-time}
\end{figure}




Figure \ref{fig:res2-model1-simd} shows the conditional effect of SIMD on receipt of social care for individiuals living in local authorities with a 96% - 99.9% match rate of SCS records to the population spine. Because SIMD was modelled as a continous variable the APE of -0.006 shown in table \ref{tab:res2-model1-APE} indicates an instantaneous effect and represents the coefficient of the slope in figure ref\ref{fig:res2-model1-simd} at the average value of SIMD (5.22, SD 2.60). As the low value indicates, after adjustment for other variables in the model, the conditinal effect of SIMD is slight and varies by 1.8% across deciles (0.029 (95%CI 0.025-0.032) in decile 1 to 0.011(95% CI 0.010-0.013) in decile 10).


Although figure \ref{fig:res2-model1-APE} is coded to include error bars plotting the 95% confidence intervals, the intervals are so small they are obscured by the effect point. As an example, in table \ref{tab:res2-model1-APE}, the largest difference between the lower and upper threshold is 0.065 (for age 95 plus). This likely reflects the large number of observations in each model which has the effect of reducing the standard error estimates. 

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_25_simd_plot.pdf}
    \caption{Subset 1. Conditional effect of simd on receipt of social care}
    \label{fig:res2-model1-simd}
\end{figure}


Figure \ref{fig:res2-model1-APE-time} charts APEs for each value of each variable (excluding local authority) across all models. There are very small increases in the APE from 2014/15 for age groups over the age of 80 and the 9 and over medicine group. All other variables have consistent APE values across models fitted to each year of data.

Figure \ref{fig:res2-model1-APELA-time} shows APEs for each councils across all models. Although there appears to be more fluctuation than seen in figure \ref{fig:res2-model1-APE-time} the scale of the y-axis is much reduced in this figure. The largest change is seen in Angus council whose APE shows a steady decline with an approximate 0.05 difference between 2011/12 and 2015/16 values.

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrrrrrr@{}}
\toprule
 & \multicolumn{1}{l}{} & \multicolumn{3}{c}{Fitted model} & \multicolumn{3}{c}{Null model} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\ \cmidrule(lr){3-5}\cmidrule(lr){6-8}
\textbf{\begin{tabular}[c]{@{}l@{}}Financial\\ Year\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}McFadden's\\ pseudo ${R^2}$\end{tabular}} & \textbf{Deviance} & \textbf{\begin{tabular}[c]{@{}r@{}}Log\\ Likelihood\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Degrees\\ of freedom\end{tabular}} & \textbf{Deviance} & \textbf{\begin{tabular}[c]{@{}r@{}}Log\\ Likelihood\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Degrees\\ of freedom\end{tabular}} & \textbf{AIC} & \textbf{BIC} \\ \midrule
\textbf{2012/13} & 0.222 & 124626.619 & -62313.309 & 223999 & 160148.076 & -80074.038 & 224053 & 124736.619 & 125304.199 \\
\textbf{2013/14} & 0.222 & 128595.553 & -64297.777 & 228726 & 165344.874 & -82672.437 & 228780 & 128705.553 & 129274.282 \\
\textbf{2014/15} & 0.233 & 131849.309 & -65924.654 & 231616 & 171893.338 & -85946.669 & 231670 & 131959.309 & 132528.728 \\
\textbf{2015/16} & 0.236 & 131443.94 & -65721.97 & 234337 & 172080.12 & -86040.06 & 234391 & 131553.94 & 132124.002 \\
\textbf{2011/12} & 0.218 & 122092.147 & -61046.073 & 218335 & 156192.796 & -78096.398 & 218389 & 122202.147 & 122768.319 \\ \bottomrule
\end{tabular}%
}
\caption{Subset 1. Goodness-of-fit}
\label{tab:res2-model1-gof}
\end{table}

Table \ref{tab:res2-model1-gof} shows the results of diagnostic tests for the models applied to each financial year of data. McFadden's pseudo $R^2$ statistic ranges from 0.252 for the model applied to 2011/12 to 0.268 for the model applied to 2015/16. This suggests the the model has an excellent fit to all years of data and gets better in more recent years.

```{r model_2, echo=FALSE, eval=FALSE}
###################################################################################################
```

\FloatBarrier
\section[Section Title, Section subtitle]{Social Care - subset 2\\ {\large Models for councils with link rate between 92\% and 95.9\%}}\label{sec:9296}

This section is a repetition of the previous section \ref{stats-rq1} with one main difference. The results are of logistic regression models run with a subset of data including individuals from the nine local authorities where a match rate of between 92% and 95.9% was achieved when linking the SCS to the population spine.

As in the previous section, descriptive statistics of the subset are first reported, followed by goodness-of-fit statistics for each model, and finally model effects reported as APEs. Again, financial year 2013/14 is used as a reference year for tables and plots where patterns are consistent across all years of data.

\subsection{Descriptive statistics}\label{subsec:res2-model2-desc}

Table \ref{tab:res2-mod2-chars1314} shows a breakdown of the cohort in financial year 2013/14. There were a total of 305.080 individuals from the nine included local authority areas in this financial year. A higher percentage of females (12.8%) compared to males (7.3%) received social care. Only 2.6% of those aged 65-69 received social care whereas 43.6% of those over the age of 95 did so. SIMD was modelled as a continuous variable. The median vale of 5 (IQR 6) was lower (more deprived) for those that received social care than for those that did not (Median 6, IQR 6). A small proportion (2.9%) of those in the lowest multimorbidity group (0-2 repeat medicines) received social care whereas over a fifth (21.2%) of those in the highest multimorbidity group received care. 


\begin{table}[]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrr@{}}
\toprule
 & \multicolumn{3}{c}{\begin{tabular}[c]{@{}c@{}}Data collected during\\ census week only\end{tabular}} & \multicolumn{2}{c}{\begin{tabular}[c]{@{}c@{}}Expanded data\\ collection for\\ telecare and \\ alarm services\end{tabular}} \\ \cmidrule(lr){2-4}\cmidrule(lr){5-6}
 & \textbf{2011/12} & \textbf{2012/13} & \textbf{2013/14} & \textbf{2014/15} & \textbf{2015/16} \\ \midrule
\textbf{Shetland Islands} & 17.8 & 16.8 & 15.9 & 14.8 & 13.9 \\
\textbf{Glasgow City} & 14.2 & 14.7 & 11.9 & 11.3 & 11.6 \\
\textbf{Na h-Eileanan Siar} & 11.6 & 10.7 & 11.6 & 11.2 & 14.6 \\
\textbf{East Renfrewshire} & 10.2 & 8.7 & 11.2 & 10.4 & 12.5 \\
\textbf{City of Edinburgh} & 9.5 & 9.1 & 9.7 & 10.5 & 9 \\
\textbf{East Dunbartonshire} & 9.6 & 9.9 & 9.7 & 10.3 & 8 \\
\textbf{Fife} & 9 & 9.3 & 9.1 & 10.5 & 11 \\
\textbf{Perth and Kinross} & 8.1 & 8.4 & 7.6 & 9.3 & 9.9 \\
\textbf{South Ayrshire} & 8.1 & 8.4 & 8.5 & 7.8 & 7.8 \\ \bottomrule
\end{tabular}%
}
\caption{Subset 2. Percentage of over 65s receiving any form of social care}
\label{tab:res2-mod2-la}
\end{table}

As there was a change in the method of data collection from 2014/15 onwards (described in section \ref{subsec:source-sc}), percentages of individuals receiving social care in each local authority is reported across all years of data in table \ref{tab:res2-mod2-la}. With the exception of financial year 2015/16, Shetland Islands council has the highest percentage of individuals over 65 receiving care in all years. It also shows a reduction in values after 2014/15 as does Glasgow City council. Other councils have more stable levels for the first three years of data and notable increases from 2014/15 onwards.  

\FloatBarrier
\section{Model results}\label{subsec:9296-AME}

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_24_AME_plot.pdf}
    \caption{Subset 2. Average partial effect plot}
    \label{fig:res2-model2-APE}
\end{figure}

Figure \ref{fig:res2-model2-APE} plots the Average Partial Effects (APEs) of each model whilst table \ref{tab:res2-mod2-APE} shows the values for the reference year 2013/14 only. Age has the largest positive effect of receiving social care with the probability approximately 40% higher for those over 95 compared to those aged 65-69 in all models (specifically 40.1% in 2013/14). Age bands 90-94, 85-89, and 80-84 also show very strong effects (~33%, ~20%, and ~12% higher probability of receiving social care across all models). 

Multimorbidity has a lesser, but still strong, effect on receipt of social care. Those receiving 9 or more repeat medicines are ~12% more likely to receive social care compared to those receiving 0-2 repeat medicines (12.7% in 2013/14). This effect is a similar magnitude as the APE of being 80-84 compared to 65-69 years of age. Medicine groups 6-8 and 3-5 have respectively ~6% and ~3% increased probability of receiving social care compared to those receiving 0-2 medicines. 

Despite large differences seen in percentages between males and females receiving social care (figure \ref{fig:res2-age-sex} and table \ref{tab:res2-mod2-chars1314}), after controlling for other variables the APE of being female compared to male is relatively small (a 3% increase in probability in 2013/14).


Figure \ref{fig:res2-model2-simd} shows the conditional effect of SIMD on receipt of social care. SIMD was modelled as a continous variable, therefore the APE of -0.005 shown in table \ref{tab:res2-mod2-APE} indicates the coefficient of the slope in figure ref\ref{fig:res2-model2-simd} at the average value of SIMD (5.76, SD 3.10). As the low value indicates, after adjustment for other variables in the model, the conditinal effect of SIMD is slight and varies by 1.1% across deciles (0.025 (95%CI 0.023-0.028) in decile 1 to 0.014(95% CI 0.012-0.015) in decile 10).  

In models from earlier years (2011/12 - 2013/14) there is a large difference between the APE of all councils compared to Shetland Islands. However, between the other councils there is much less variation in APEs (e.g in 2013/14 there is a  2.9% difference in the probability of receiving social care between the extremes of East Renfrewshire and South Ayrshire). In later models (2014/15 & 2015/16) the variation in APEs has a more linear trend and ranges by ~10% difference in the probability of receiving social care. 

As in previous models, confidence intervals are very tight and difficult to visualise. Again, this is likely due to the large number of observations included in the models and the resultant reduction of standard errors. 

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-2/sc_model_24_council_APE_plot.pdf}
    \caption{Subset 2. Local authority APE across models}
    \label{fig:res2-model2-APE-time}
\end{figure}

Figure \ref{fig:res2-model2-APE-time} charts APEs for each value of each variable (excluding local authority) across all models. There are increases in the APE from 2014/15 for age groups over the age of 80 and the 9 and over medicine group. This is more marked for older age groups. For example there is a ~8% increase in probability of receiving social care for over 95s from 2014/15 onwards. All other variables have consistent APE values across models fitted to each year of data.


Figure \ref{fig:res2-model2-APELA-time} shows APEs for each council across all models. The most notable trend is the decrease in APEs across all councils in 2014/14 compared to Shetland Islands. This size of this decrease is larger in some councils (e.g. Perth & Kinross ~ 0.07) compared to others (e.g. Glasgow City ~ 0.02). Again the difference in scale compared to figure \ref{fig:res2-model2-APE-time} should be noted. The relative size of the differences in APE is small in comparison to, for example, the difference between the APE of being 95 plus in 2013/14 and 2014/15 seen in the previous figure. 

\begin{table}[]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrrrrrr@{}}
\toprule
 & \multicolumn{1}{l}{} & \multicolumn{3}{c}{Fitted model} & \multicolumn{3}{c}{Null model} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\ \cmidrule(lr){3-5}\cmidrule(lr){6-8}
\textbf{\begin{tabular}[c]{@{}l@{}}Financial\\ Year\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}McFadden's\\ pseudo ${R^2}$\end{tabular}} & \textbf{Deviance} & \textbf{\begin{tabular}[c]{@{}r@{}}Log\\ Likelihood\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Degrees\\ of freedom\end{tabular}} & \textbf{Deviance} & \textbf{\begin{tabular}[c]{@{}r@{}}Log\\ Likelihood\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Degrees\\ of freedom\end{tabular}} & \textbf{AIC} & \textbf{BIC} \\ \midrule
\textbf{2011/12} & 0.196 & 159534.227 & -79767.114 & 293867 & 198430.468 & -99215.234 & 293921 & 159644.227 & 160226.736 \\
\textbf{2012/13} & 0.198 & 162331.401 & -81165.7 & 299514 & 202476.6 & -101238.3 & 299568 & 162441.401 & 163024.956 \\
\textbf{2013/14} & 0.199 & 159148.11 & -79574.055 & 305011 & 198718.545 & -99359.2725 & 305065 & 159258.11 & 159842.666 \\
\textbf{2014/15} & 0.204 & 164081.461 & -82040.731 & 308392 & 206217.188 & -103108.594 & 308446 & 164191.461 & 164776.623 \\
\textbf{2015/16} & 0.209 & 163509.29 & -81754.645 & 310949 & 206829.494 & -103414.747 & 311003 & 163619.29 & 164204.906 \\ \bottomrule
\end{tabular}%
}
\caption{Subset 2. Goodness-of-fit}
\label{tab:res2-mod2-gof}
\end{table}

Table \ref{tab:res2-mod2-gof} shows the results of diagnostic tests for the models applied to each financial year of data. McFadden's pseudo $R^2$ statistic ranges from 0.195 for the model applied to 2011/12 to 0.219 for the model applied to 2015/16. The models applied to the most recent two years of data have a good fit, whereas the previous three models are only marginally below the threshold of 0.2 generally regarded as representing an excellent fit.

\FloatBarrier
\section{Unplanned admission}\label{sec:res3-usc}

This section reports results of logistic regression models detailed in section \ref{stats-rq2}. The outcome variable in these models is acute admission to hospital in the six months after the end of each financial year. 

The section follows the same format as presented in the previous section with the outcome variable being the only difference in analyses.  

\FloatBarrier
\subsection{Descriptive statistics}\label{subsec:res3-adm-desc}


\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{threeparttable}
\begin{tabular}{@{}llrrrr@{}}
\toprule
\textbf{Variable} & \textbf{Value} & \textbf{\begin{tabular}[c]{@{}r@{}}No unplanned\\ admission\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}At least one\\ unplanned admission\end{tabular}} & \textbf{Total} & \textbf{p-value} \\ \midrule
\textbf{Sex} & Male & 362216 (90.9) & 36145 (9.1) & 398361 & 0.221 \\
\textbf{} & Female & 444060 (91.0) & 43910 (9.0) & 487970 &  \\
\textbf{Age group} & 65-69 & 264091 (94.8) & 14522 (5.2) & 278613 & \textless{}0.001 \\
\textbf{} & 70-74 & 196494 (93.2) & 14418 (6.8) & 210912 &  \\
\textbf{} & 75-79 & 152457 (90.4) & 16208 (9.6) & 168665 &  \\
\textbf{} & 80-84 & 106369 (86.9) & 15983 (13.1) & 122352 &  \\
\textbf{} & 85-89 & 58183 (83.1) & 11831 (16.9) & 70014 &  \\
\textbf{} & 90-94 & 22701 (80.2) & 5610 (19.8) & 28311 &  \\
\textbf{} & 95 plus & 5981 (80.1) & 1483 (19.9) & 7464 &  \\
\textbf{SIMD decile} & 1 - most deprived & 63364 (86.9) & 9527 (13.1) & 72891 & \textless{}0.001 \\
\textbf{} & 2 & 71876 (88.6) & 9247 (11.4) & 81123 &  \\
\textbf{} & 3 & 80768 (89.4) & 9602 (10.6) & 90370 &  \\
\textbf{} & 4 & 83216 (90.3) & 8947 (9.7) & 92163 &  \\
\textbf{} & 5 & 85840 (91.0) & 8456 (9.0) & 94296 &  \\
\textbf{} & 6 & 87908 (91.6) & 8027 (8.4) & 95935 &  \\
\textbf{} & 7 & 84892 (92.2) & 7215 (7.8) & 92107 &  \\
\textbf{} & 8 & 82592 (92.6) & 6602 (7.4) & 89194 &  \\
\textbf{} & 9 & 83433 (92.8) & 6519 (7.2) & 89952 &  \\
\textbf{} & 10 - most affluent & 82387 (93.3) & 5913 (6.7) & 88300 &  \\
\textbf{Repeat medicines group} & 0-2 & 233426 (96.4) & 8617 (3.6) & 242043 & \textless{}0.001 \\
\textbf{} & 3-5 & 236155 (93.7) & 15852 (6.3) & 252007 &  \\
\textbf{} & 6-8 & 181810 (89.8) & 20653 (10.2) & 202463 &  \\
\textbf{} & 9+ & 154885 (81.6) & 34933 (18.4) & 189818 &  \\
\textbf{Any form of social care} & No Social Care & 729724 (92.8) & 56233 (7.2) & 785957 & \textless{}0.001 \\
\textbf{} & Social Care & 76552 (76.3) & 23822 (23.7) & 100374 &  \\
\textbf{Any USC episode in previous year} & No USC & 601891 (94.5) & 34814 (5.5) & 636705 & \textless{}0.001 \\
\textbf{} & USC & 204385 (81.9) & 45241 (18.1) & 249626 &  \\
\textbf{Residence} & Not Care Home & 777019 (91.2) & 75348 (8.8) & 852367 & \textless{}0.001 \\
\textbf{} & Care Home & 23535 (83.7) & 4592 (16.3) & 28127 &  \\ \bottomrule
\end{tabular}%
\begin{tablenotes}
\item \footnotesize{Total N used in model = 868,947}
\item \footnotesize{All values n(\%) unless otherwise stated} 
\item \footnotesize{\emph{Row}-wise percentages within grouped variables}
\end{tablenotes}
\end{threeparttable}%
}
\caption{Characteristics of those with/without any unplanned hospital admission in the six months following end of 2015/16}
\label{tab:res3adm-chars1516}
\end{table}


Table \ref{tab:res3adm-chars1516} shows a cross table of the cohort in 2015/16 with regard to the outcome variable. As seen in the previous section, there were a total of 868,947 individuals included in the model for this financial year. There was little difference between the percentage of males (9.2%) and females (9.1%) that were admitted to hospital during the outcome period and the difference between these groups was not statistically significant. A majority (55%) of those that died during the outcome period were admitted to hospital. An increase in the percentage of individuals being admitted to hospital is seen through age groups from 5.1% of those aged 65-69 to 20% of those aged 90-94 with a slight reduction in those aged over 95 (19.7%). Increases are also seen through multimorbidity groups from 3.6% for those prescribed 0-2 repeat medicines to 18.4% for those prescribed nine or more repeat medicines. Higher percentages of the population using social care (24.1%), unscheduled care in the year prior to the outcome period (18.7%), and those resident in care home (17.1%) had an admission in the outcome period compared to those not using these services (7.3%, 5.6%, and 8.9% respectively). The median value of SIMD (5, IQR 4) was lower than that for those with no admission (6, IQR 5). 

\FloatBarrier
\subsection{Model fit}

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-3/adm_model_3_or_plot.pdf}
    \caption{Unplanned hospital admission. Odds ratios}
    \label{fig:res3-adm-APE}
\end{figure}


\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-3/adm_model_3_AME_plot.pdf}
    \caption{Unplanned hospital admissions. Average Partial Effects}
    \label{fig:res3-adm-APE}
\end{figure}

Figure \ref{fig:res3-adm-APE} plots the APEs of each variable in each model and table \ref{tab:res3-adm-ape1516} shows the values of APEs for financial year 2013/14. As reported in models where any USC use is the outcome variable in the previous section, dying during the outcome period has the largest APE. This magnitude of this effect is, however, lower than seen in the previous section (34.6% increase in probability of admission compared to almost 60% increase in any USC use). High multimorbidity (nine or more repeat prescriptions) has the next highest positive APE, In 2013/14 this group had a 7.7% increase in probability in being admitted to hospital compared to those from the lowest group (0-2 repeat medicines). Those with 6-8 and 3-5 repeat medicines had increases in probability of admission of 3.5% and 1.6% respectively. Use of any form of unscheduled care in the year prior to the outcome period or receipt of social care during the SCS census week was associated with increases in the probability of admission to hospital in 2013/14 of 6.7% compared to those with no prior USC use and 6.0% for those with no social care receipt respectively. The was an incremental increase in APE through age groups from those aged 70-74 (0.7%) to those aged 90-94 (6.0%) compared to those age 65-69. Those aged 95 or above had a lower APE (5.4%) than their immediately younger age group. 

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
\begin{threeparttable}
\begin{tabular}{@{}lrrrrr@{}}
\toprule
\textbf{Variable} & \textbf{\begin{tabular}[c]{@{}r@{}}Average\\ Partial Effect\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Standard\\ Error\end{tabular}} & \textbf{p-value} & \textbf{\begin{tabular}[c]{@{}r@{}}Lower 95\%\\ CI\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Upper 95\%\\ CI\end{tabular}} \\ \midrule
\textbf{Meds: 9+} & 0.083 & 0.001 & \textless{}0.001 & 0.081 & 0.085 \\
\textbf{Age: 90-94} & 0.075 & 0.003 & \textless{}0.001 & 0.07 & 0.08 \\
\textbf{USC} & 0.073 & 0.001 & \textless{}0.001 & 0.072 & 0.075 \\
\textbf{Social Care} & 0.065 & 0.001 & \textless{}0.001 & 0.063 & 0.067 \\
\textbf{Age: 95 Plus} & 0.06 & 0.005 & \textless{}0.001 & 0.051 & 0.07 \\
\textbf{Age: 85-89} & 0.057 & 0.001 & \textless{}0.001 & 0.054 & 0.06 \\
\textbf{Age: 80-84} & 0.037 & 0.001 & \textless{}0.001 & 0.034 & 0.039 \\
\textbf{Meds: 6-8} & 0.037 & 0.001 & \textless{}0.001 & 0.035 & 0.039 \\
\textbf{Age: 75-79} & 0.021 & 0.001 & \textless{}0.001 & 0.019 & 0.023 \\
\textbf{Meds: 3-5} & 0.015 & 0.001 & \textless{}0.001 & 0.014 & 0.017 \\
\textbf{Age: 70-74} & 0.007 & 0.001 & \textless{}0.001 & 0.005 & 0.009 \\
\textbf{SIMD Decile:} & -0.003 & 0 & \textless{}0.001 & -0.003 & -0.003 \\
\textbf{Care Home} & -0.008 & 0.001 & \textless{}0.001 & -0.011 & -0.006 \\
\textbf{Female} & -0.016 & 0.001 & \textless{}0.001 & -0.017 & -0.015 \\ \bottomrule
\end{tabular}%
\begin{tablenotes}
\item \footnotesize{Reference group for age: 65-69}
\item \footnotesize{Reference group for meds: 0-2 repeat medicines}
\end{tablenotes}
\end{threeparttable}%
}
\caption{Unplanned hospital admission. Average Partial Effects. 2015/16}
\label{tab:res3-adm-ape1516}
\end{table}

There is a negative effect on admission to hospital for females compared to males with a difference in probability of 1.3%. This effect size is larger than seen in models where any form of USC was the outcome variable (section \ref{subsec:res3-usc-ape}). Living in a care home also had a negative effect on admission to hospital after controlling for other variables in the model. The APE in 2013/14 was -0.011 indicating a 1.1% decrease in the probability of admission to hospital for those living in a care home. This value increases over time and the most recent model shows a 3% decrease in probability of admission. 

Figure \ref{fig:res3-adm-simd} shows the conditional effect of SIMD on admission to hospital during the six months following financial year 2013/14. As SIMD was modelled as a continuous variable the APE of -0.003 represents the coefficient of the slope in figure ref at the average value of SIMD (5.6, SD 2.8). After adjusting for other variables the conditional effect of SIMD varies by only 1% across deciles (0.034 (95% CI 0.032-0.036) in decile 1 to 0.024(95% CI 0.023-0.025) in decile 10).

\begin{figure}[h]
  \centering
    \includegraphics{figures/chapter-results-3/adm_model_3_simd_plot.pdf}
    \caption{Conditional effect of SIMD on any unplanned hospital admission}
    \label{fig:res3-adm-simd}
\end{figure}


\begin{table}[]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrrrrrr@{}}
\toprule
 & \multicolumn{1}{l}{} & \multicolumn{3}{c}{Fitted model} & \multicolumn{3}{c}{Null model} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\ \cmidrule(lr){3-5}\cmidrule(lr){6-8}
\textbf{\begin{tabular}[c]{@{}l@{}}Financial\\ Year\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}McFadden's\\ pseudo ${R^2}$\end{tabular}} & \textbf{Deviance} & \textbf{\begin{tabular}[c]{@{}r@{}}Log\\ Likelihood\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Degrees\\ of freedom\end{tabular}} & \textbf{Deviance} & \textbf{\begin{tabular}[c]{@{}r@{}}Log\\ Likelihood\end{tabular}} & \textbf{\begin{tabular}[c]{@{}r@{}}Degrees\\ of freedome\end{tabular}} & \textbf{AIC} & \textbf{BIC} \\ \midrule
\textbf{2011/12} & 0.113 & 447862.4 & -223931.2 & 821187.0 & 504865.0 & -252432.5 & 821234.0 & 447958.4 & 448516.1 \\
\textbf{2012/13} & 0.112 & 454527.8 & -227263.9 & 841129.0 & 512028.3 & -256014.2 & 841176.0 & 454623.8 & 455182.7 \\
\textbf{2013/14} & 0.114 & 466492.6 & -233246.3 & 858910.0 & 526463.8 & -263231.9 & 858957.0 & 466588.6 & 467148.4 \\
\textbf{2014/15} & 0.111 & 471115.4 & -235557.7 & 870332.0 & 530199.5 & -265099.8 & 870379.0 & 471211.4 & 471771.8 \\
\textbf{2015/16} & 0.113 & 475554.4 & -237777.2 & 880446.0 & 535977.2 & -267988.6 & 880493.0 & 475650.4 & 476211.4 \\ \bottomrule
\end{tabular}%
}
\caption{Unplanned hospital admission. Goodness-of-fit}
\label{res3-adm-gof}
\end{table}

Table \ref{res3-adm-gof} shows the results of diagnostic tests for each model. The value of McFadden's $R^2$ decreases slightly over time from 0.147 to 0.142. Although values are lower than seen for social care models fitted in the previous chapter, they are slightly higher than found for models using any USC use reported in the previous section. 


