---
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    citation_package: natbib
    keep_tex: false
documentclass: report
classoption: twoside
bibliography: thesis
csl:  harvard-university-of-the-west-of-scotland.csl
urlcolor: cyan
linkcolor: blue
fontsize: 12pt
geometry: margin = 1.2in
header-includes:
- \usepackage[section]{placeins}
- \usepackage{caption}
- \usepackage{fancyhdr}
- \usepackage{setspace}
- \usepackage{chngcntr}
- \usepackage[export]{adjustbox}
- \usepackage{microtype}
- \usepackage{booktabs}
- \usepackage{graphicx}
- \usepackage{epigraph}
- \usepackage{pdfpages}
- \usepackage[inline]{enumitem}
- \usepackage[flushleft]{threeparttable}
- \usepackage{hyperref}
- \usepackage[toc,page]{appendix}
- \usepackage[utf8]{inputenc}
- \onehalfspacing
- \colorlet{Blue}{blue} #Sort out an issue with xcolor package(not explicitly loaded here!)
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path = 'figures/', echo = FALSE,
                      warning = FALSE, message = FALSE)
```

\pagenumbering{gobble}

```{r child = 'titlepage.Rmd'}
```

\newpage
\pagestyle{fancy}

\fancyhead[LE,RO]{}
\fancyhead[LO,RE]{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\pagenumbering{roman}

\section*{Abstract}
\addcontentsline{toc}{section}{Abstract}

```{r child = 'abstract.Rmd'}
```

\newpage
\fancyhead[CO,CE]{Table of Contents}
\setcounter{tocdepth}{2}
\tableofcontents

\newpage
\fancyhead[CO,CE]{List of Tables}
\addcontentsline{toc}{section}{List of Tables}
\listoftables

\newpage
\fancyhead[CO,CE]{List of Figures}
\addcontentsline{toc}{section}{List of Figures}
\listoffigures

\newpage
\section*{Acknowledgements}
\fancyhead[CO,CE]{}
\renewcommand{\headrulewidth}{0pt}
\addcontentsline{toc}{section}{Acknowledgements}

```{r child = 'acknowledgements.Rmd'}
```

\newpage
\section*{Declaration}
\fancyhead[CO,CE]{}
\renewcommand{\headrulewidth}{0pt}
\addcontentsline{toc}{section}{Declaration}

```{r child = 'declaration.Rmd'}
```


\newpage
\pagenumbering{arabic}

\newpage
\fancyhf{}
\fancyhead[CE,CO]{\leftmark}
\fancyfoot[LE,LO]{\rightmark}
\fancyfoot[RE,RO]{Page \thepage}
\renewcommand{\headrulewidth}{0pt}

```{r child = 'intro.Rmd'}
```

\FloatBarrier
\newpage


```{r child = 'lit_review.Rmd'}
```

\FloatBarrier
\newpage

```{r child = 'renfrew.Rmd'}
```


\FloatBarrier
\newpage

```{r child = 'data.Rmd'}
```


\FloatBarrier
\newpage

```{r child = 'methods.Rmd'}
```

\FloatBarrier
\newpage

```{r child = 'results.Rmd'}
```

\FloatBarrier
\newpage

```{r child = 'conclusion.Rmd'}
```


\FloatBarrier
\newpage


```{r child = 'appendix_1.Rmd'}
```

\FloatBarrier
\clearpage

\addcontentsline{toc}{chapter}{Bibliography}
