---
title: "Create plot for SC model percentages"
author: "David Henderson"
date: "17 March 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Intro

Each sc model has a table with the percentage of individuals receiving social care in each local authority over time. I want to create a plot of this table. 

##Packages

```{r libs, message=FALSE}
library(tidyverse)
library(magrittr)
library(ggthemes)
library(ggrepel)
library(ggforce)
library(here)
here()
theme_set(theme_minimal(base_size = 16))
```

##Data

I have downloaded the csv from the html document `62_social_care_model_25` in the folder "thesis_rmd/figures/chapter-results-2".

```{r load, message=FALSE, fig.width = 5.25, fig.height = 7}
sc_25 <- read_csv(here("data/assets/sc_model_25_tab_2.csv"))
simd <- read_csv(here("data/assets/simd_la.csv")) 
```


#Tidy and plot

First thing is to change the values in the table so that they reflect the difference in the percentage of over 65s receiving social care in each council area. I then drop Falkirk and plot the the results. This plot can be used to compare against the partial effects after accounting for age, sex, multimorbidity status etc.  

```{r tidy_up, fig.width=10.67, fig.height=8}
sc_25 %>%
  select(-X1) %>% 
  mutate(`2011/12` = `2011/12` - first(`2011/12`),
         `2012/13` = `2012/13` - first(`2012/13`),
         `2013/14` = `2013/14` - first(`2013/14`),
         `2014/15` = `2014/15` - first(`2014/15`),
         `2015/16` = `2015/16` - first(`2015/16`)) %>% 
  gather(year, pct, `2011/12`:`2015/16`) %>% 
  arrange(council) %>% 
  mutate(label = if_else(year == max(year),
                         council, NA_character_)) %>% 
  filter(council != "Falkirk") %>% 
  ggplot(aes(year, pct,
             group = council,
             colour = council)) +
  geom_point(size = 2) +
  geom_path() +
  geom_hline(yintercept = 0, colour = "#555555") +
  geom_vline(xintercept = 3.5, linetype = "longdash", colour = "grey30") +
  geom_label_repel(aes(label = label), nudge_x = 2) +
  geom_label(aes(x = "2011/12", y = -0.001),
             label = "Reference group: Falkirk",
             colour = "#555555",
             hjust = 0) +
  geom_text(aes(3.55, y = 0.9),
            label = "Change in data collection after 2013/14",
            colour = "#AAAAAA",
            hjust = 0) +
  scale_colour_ptol(guide = "none") +
  scale_y_continuous(limits = c(-8, 1.5),
                     breaks = scales::pretty_breaks()) +
  scale_x_discrete(expand = expand_scale(add = c(0.1, 0.9))) +
  labs(x = "",
       y = "% difference from Falkirk council, by year",
       title = "Percentage difference in over 65s receiving social care",
       subtitle = "compared to Falkirk council") -> sc_25_pct_plot
sc_25_pct_plot
  
```

#Save

```{r save_1}
ggsave(here("figures/chapter-results-2/sc_25_pct_plot.png"),
       width = 10.67, height = 8, dpi = 600)
```

#SC model_24

Now repeat for the model with lower linkage rates

##Data

I have downloaded the csv from the html document `61_social_care_model_24` in the folder "thesis_rmd/figures/chapter-results-2".

```{r load, message=FALSE}
sc_24 <- read_csv(here("data/assets/sc_model_24_tab_2.csv"))
```

#Alternative plot

```{r sc, fig.width=9, fig.height=6.75}
sc_25 %>%
  select(council, `2015/16`) %>% 
  rename(pct = `2015/16`) %>% 
  filter(council != "Falkirk") %>% 
  mutate(ape = c(-0.016, -0.01, -0.012, -0.033, -0.01, -0.029,
                 -0.05, -0.041)) %>%
  left_join(., simd) %>% 
  ggplot(aes(simd_wt, ape, colour = council)) +
  geom_point() +
  geom_label_repel(aes(label = council), 
                   nudge_x = 0.01) +
  scale_colour_ptol(guide = "none") +
  scale_y_continuous(limits = c(-0.08, 0)) +
  scale_x_continuous(limits = c(0, 15)) +
  labs(x = "weighted average simd %",
       y = "Average Partial Effect of receiving social care",
       title = "Comparison of percentage vs partial effect receiving social care",
       subtitle = "2015/16") 
```


```{r av_simd, fig.width=9, fig.height=6.75}
simd %>%  
  group_by(council) %>% 
  mutate(wt_pct = (simd * frq),
         simd_wt = round(sum(wt_pct/55)*100,3)) %>% 
  select(council, simd_wt) %>% 
  distinct(council, .keep_all = TRUE) %>% 
  arrange(-simd_wt) -> simd
simd %>% 
  ggplot(aes(reorder(council, simd_wt), simd_wt)) +
  geom_col() +
  coord_flip() +
  labs(x = "")
```



#Tidy and plot

First thing is to change the values in the table so that they reflect the difference in the percentage of over 65s receiving social care in each council area. I then drop Falkirk and plot the the results. This plot can be used to compare against the partial effects after accounting for age, sex, multimorbidity status etc.  

```{r tidy_up, fig.width=10.67, fig.height=8}
sc_24 %>%
  select(-X1) %>% 
  mutate(`2011/12` = `2011/12` - first(`2011/12`),
         `2012/13` = `2012/13` - first(`2012/13`),
         `2013/14` = `2013/14` - first(`2013/14`),
         `2014/15` = `2014/15` - first(`2014/15`),
         `2015/16` = `2015/16` - first(`2015/16`)) %>% 
  gather(year, pct, `2011/12`:`2015/16`) %>% 
  arrange(council) %>% 
  mutate(label = if_else(year == max(year),
                         council, NA_character_)) %>% 
  filter(council != "Shetland Islands") %>% 
  ggplot(aes(year, pct,
             group = council,
             colour = council)) +
  geom_point(size = 2) +
  geom_path() +
  geom_hline(yintercept = 0, colour = "#555555") +
  geom_vline(xintercept = 3.5, linetype = "longdash", colour = "grey30") +
  geom_label_repel(aes(label = label), nudge_x = 2) +
  geom_label(aes(x = "2011/12", y = -0.001),
             label = "Reference group: Shetland Islands",
             colour = "#555555",
             hjust = 0) +
  geom_text(aes(3.55, y = 0.9),
            label = "Change in data collection after 2013/14",
            colour = "#AAAAAA",
            hjust = 0) +
  scale_colour_ptol(guide = "none") +
  scale_y_continuous(limits = c(-12, 1.5),
                     breaks = scales::pretty_breaks()) +
  scale_x_discrete(expand = expand_scale(add = c(0.1, 0.9))) +
  labs(x = "",
       y = "% difference from Shetland council, by year",
       title = "Percentage difference in over 65s receiving social care",
       subtitle = "compared to Shetland Islands council") -> sc_24_pct_plot
sc_24_pct_plot
  
```

#Save

```{r save_1}
ggsave(here("figures/chapter-results-2/sc_24_pct_plot.png"),
       width = 10.67, height = 8, dpi = 600)
```




