---
title: "OECD data"
output: html_notebook
---

#Introduction

A short script to import data from the OECD and create a visualisation of variation in long-term care provision across countries.

#Preparation
##Load libraries

```{r lib, warning=FALSE, message=FALSE}
library(tidyverse)
```

##Import Data
I have downloaded the data from the OECD site as I have to request the format (i.e. I can't direct import to R). The data is avaiable from this link <http://stats.oecd.org/viewhtml.aspx?datasetcode=HEALTH_LTCR&lang=en>. I have saved it as "raw_oecd_data.csv" in my working directory. Data was extracted on 22nd August 2017 at 09:30hrs GMT., 

```{r import}
oecd_raw <- read_csv(file = "data/raw_oecd_data.csv")
```

Ok, so the raw data is imported and we can see all variables have been imported as character strings. Plenty wrangling to do but first I am going to save this object in its most raw format so we have a copy of that.

```{r save_raw, echo=FALSE}
save(oecd_raw, file = "data/produced_data/oecd_raw.rds")
```

#Wrangle
* Tidy variable types, names etc.

#Explore
* Missing
* Assertions

#Visualise
*re-create OECD plot

#Scotland data

##Import data

##Wrangle
* tidy

##Explore
* Missing
* Assertions

##Summarise
* calculate Scotland average (previously done by LA)

#Add Scotland to OECD graph
